% Browser-Native Spiking Neural Networks in Contracted Physics
% NeurIPS 2027 Workshop on Neuromorphic Computing / Embodied AI
% Paper ID: P1-SNN (SNN compute path under contract in Program 1: Trust-by-Construction)
%
% Build: pdflatex paper-2-snn-neurips.tex
%        bibtex paper-2-snn-neurips
%        pdflatex paper-2-snn-neurips.tex (x2)

\documentclass[10pt]{article}

% --- NeurIPS-style formatting ---
% Guarded load: matches the USENIX .sty pattern used by papers 1 and 4
% (see paper-audit-matrix.md venue-fit note, 2026-04-19). When the vendor
% neurips_<year>.sty is present on submission, this picks it up cleanly.
% When absent (local compile without the style package), we fall back to a
% reasonable default (1in margins, Times). Bundle-time step on submission:
% ship neurips_2027.sty alongside the .tex.
\IfFileExists{neurips_2027.sty}{%
  \usepackage[final]{neurips_2027}%
}{%
  \typeout{Skipping neurips_2027: vendor sty absent; local compile uses default article class.}%
  \usepackage[margin=1in]{geometry}%
  \usepackage{times}%
  \setlength{\parskip}{0.5em}%
}

\usepackage[utf8]{inputenc}
\usepackage[T1]{fontenc}
\usepackage{hyperref}
\usepackage{url}
\usepackage{booktabs}
\usepackage{amsfonts}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{amsthm}
\usepackage{enumitem}  % for [leftmargin=*,topsep=2pt] enumerate options
\newtheorem{theorem}{Theorem}
\newtheorem{lemma}{Lemma}
\newtheorem{definition}{Definition}
\newtheorem{property}{Property}
\newtheorem{corollary}{Corollary}
\usepackage{nicefrac}
\usepackage{microtype}
\usepackage{graphicx}
\usepackage{xcolor}
\usepackage{algorithm}
\usepackage{algorithmic}
\usepackage{listings}
\usepackage{subcaption}
\usepackage{tikz}
\usetikzlibrary{arrows.meta, positioning, shapes.geometric, calc}

% --- Custom commands ---
\newcommand{\todo}[1]{\textcolor{red}{\textbf{[TODO: #1]}}}
\newcommand{\snnwebgpu}{\textsc{SNN-WebGPU}}
\newcommand{\holoscript}{\textsc{HoloScript}}
\newcommand{\cael}{\textsc{CAEL}}
\newcommand{\relu}{\operatorname{ReLU}}
\newcommand{\tropplus}{\oplus}
\newcommand{\tropmult}{\otimes}
\DeclareMathOperator*{\argmax}{arg\,max}

% Lotus petal data contract — \measuredFrom{<artifact>} marks values that come
% from a runnable artifact in the repo. Renders nothing in the PDF; grep target
% for scripts/build-lotus.mjs (research/2026-04-26_lotus-petal-data-contract.md §3+§4).
\providecommand{\measuredFrom}[1]{}
\providecommand{\measuredAt}[1]{}

% --- Listings style for WGSL ---
\lstdefinelanguage{WGSL}{
  keywords={fn, let, var, if, else, for, return, struct, const, true, false},
  morekeywords={f32, u32, i32, vec3, array, compute, workgroup_size, builtin, group, binding, uniform, storage, read, read_write},
  sensitive=true,
  morecomment=[l]{//},
  morestring=[b]",
}
\lstset{
  language=WGSL,
  basicstyle=\ttfamily\scriptsize,
  keywordstyle=\color{blue}\bfseries,
  commentstyle=\color{gray}\itshape,
  numbers=left,
  numberstyle=\tiny\color{gray},
  breaklines=true,
  frame=single,
  captionpos=b,
}

% ============================================================================
\title{Browser-Native Spiking Neural Networks in Contracted Physics:\\
       WebGPU Compute Shaders for Embodied Neuromorphic Cognition}

\author{
  Joseph Krzywoszyja \\
  Infinitus Systems \\
  \texttt{josep@holoscript.net} \\
}

\begin{document}

\maketitle

% ============================================================================
\begin{abstract}
We present \snnwebgpu{} as the first executable Tier~1 of an \emph{inverted
runtime stack}: neuromorphic computation is the primary executor, not an
accelerator bolted onto a CPU-centric engine, and every spike decision is backed
by a cryptographic determinism receipt that makes NN-primary runtime stacks
architecturally honest.
The system implements Leaky Integrate-and-Fire (LIF) neurons, Spike-Timing-Dependent
Plasticity (STDP) and multi-layer SNN architectures entirely in WGSL compute
shaders dispatched from a standard web browser via the WebGPU API, requiring
zero installation and no server-side computation.
We pair a small, fast SNN proposer with a CPU-side verifier
(SimulationContract) generalizing the 2026 speculative-decoding pattern from
token-stream to operation-stream; on neuromorphic-class hardware this inversion
delivers single-digit milliwatt steady-state power---three orders of magnitude
below GPU-accelerated LLM inference.
We establish a formal bridge between SNN rate coding and classical ReLU activations
via tropical algebra ($\max(0,x) = x \tropplus 0$ in the max-plus semiring),
enabling principled interoperability between spiking and non-spiking networks.
Additionally, we introduce GPU-accelerated tropical min-plus GEMM for shortest-path
computation on neural connectivity graphs, achieving $20.64\times$ speedup over
CPU at $N=1024$ nodes.
We evaluate across consumer integrated and discrete GPUs using the same WGSL
kernel and benchmark harness. The dated benchmark snapshot
\texttt{research/benchmark-results-2026-04-12.md} reports multi-billion
neuron-timesteps/second on discrete hardware and materially lower---but still
interactive---throughput on integrated hardware, with no shader changes across
devices. We treat those figures as dated artifact numbers and keep the
cross-hardware claim scoped to what is reproducibly logged in that snapshot.
The \cael{} hash-chain provenance adds $3.40\;\mu$s per entry median ($3.74\;\mu$s p99,
99{,}997-entry verify harness; Table~\ref{tab:cael}; cross-paper anchor \texttt{research/benchmark-canonical-run.md}) across configurations.
In an embodied navigation task (harness:
\texttt{paper-snn-navigation.test.ts}, 2026-04-17), the 128-neuron agent
reaches the goal while maintaining sub-millisecond median tick time on the
harness host and producing a fully hash-chained provenance record with
100\% integrity.
Every perception--cognition--action cycle is committed to a hash-chain trace,
producing the first \emph{compute substrate} to emit auditable, tamper-evident,
replayable traces of neuromorphic decisions at hardware throughput in an
embodied simulation.
% AUDIT-2026-04-20: Scope-sharpened to disambiguate from Paper 0c (contract-framework novelty).
The \texttt{@holoscript/snn-webgpu} package is validated by a dedicated Vitest
suite in the repository (full-package run reference: 2026-04-17).
Headline GPU throughputs and tropical GEMM figures above reference the dated snapshot
\texttt{research/benchmark-results-2026-04-12.md}; interactive-scale neuron counts
(max verified discrete run: RTX~3060 Laptop GPU, \texttt{strictAdapter=true},
peak $N=1{,}048{,}576$ at $5{,}890.876$ M neurons/s;
artifact \texttt{HoloScript/.bench-logs/paper-2-lif-throughput-rtx3060.json},
generated 2026-05-06 on commit \texttt{d0a990964},
driver 591.74) are measured on
commodity integrated and discrete GPUs and run in any WebGPU-capable browser.
\end{abstract}

% ============================================================================
\section{Introduction}
\label{sec:intro}

Spiking neural networks (SNNs) offer a biologically plausible model of neural computation
in which information is encoded in the precise timing of discrete action potentials (spikes)
rather than continuous activations~\cite{maass1997networks, gerstner2002spiking}.
When coupled with physics simulations, SNNs enable \emph{embodied neuromorphic cognition}:
agents that perceive physical sensor readings, process them through biologically realistic
neural dynamics, and act upon the simulation in a closed loop.

This paradigm has been explored in robotics~\cite{tang2024braincog} and neuroscience
simulation~\cite{gewaltig2007nest}, but existing systems share a critical limitation:
they require high-performance computing clusters (NEST+Gazebo~\cite{gewaltig2007nest}),
custom neuromorphic chips (Intel Loihi~\cite{davies2018loihi}, IBM TrueNorth~\cite{merolla2014truenorth}),
or GPU-accelerated server infrastructure (BrainCog~\cite{tang2024braincog}).
No existing system delivers embodied SNN cognition in a standard web browser
with zero installation.

However, browser-native deployment is not merely a convenience feature; it is a
consequence of a deeper architectural inversion we term \emph{NN-primary,
CPU-verifier backup}. In this paradigm, spiking neural networks serve as the
primary runtime executor (Tier~1), LLM speculation serves as a warm-path
verifiable proposer (Tier~2), and CPU-side AST/typecheck/CURE verification
serves as the cold-path referee (Tier~3). The browser is the deployment
surface because WebGPU and WebNN are now production-ready across all major
engines (January 2026), making the NN-primary stack portable to commodity
hardware without datacenter dependency. On neuromorphic-class silicon this
inversion achieves single-digit milliwatt steady-state power, three orders of
magnitude below GPU-accelerated LLM inference, aligning with always-on embodied
agent workloads.

We address this gap with three contributions:

\begin{enumerate}
\item \textbf{Browser-native SNN on WebGPU as Tier~1 of an inverted runtime stack.}
We implement LIF neurons, STDP learning, multi-layer networks, and spike encoding/decoding
entirely in WGSL compute shaders, dispatched via the WebGPU API.
The system runs in Chrome, Edge, and Firefox---no plugins, no server, no compilation.
Every spike decision is committed to a \cael{} trace with cryptographic hash chaining,
producing the first tamper-evident audit trail of neuromorphic decision-making in
physics simulation and yielding a \emph{determinism receipt} that makes NN-primary
execution architecturally honest.

\item \textbf{Tropical algebra bridge between SNN and ReLU.}
We formalize the observation that LIF firing rate approximates $\relu(x)$ and show this
is not a coincidence but a consequence of the tropical max-plus semiring:
$\max(0, x) = x \tropplus 0$ where $\tropplus = \max$.
This bridge enables principled conversion between spiking and non-spiking representations.

\item \textbf{Three-tier dispatch architecture and energy analysis.}
We generalize the speculative-decoding pattern from token-stream to operation-stream,
laying out the full Tier~1 (SNN hot path), Tier~2 (LLM speculative warm path with
CPU verifier), and Tier~3 (CPU cold-path referee) architecture, and quantify the
energy gap between Tier~1 on neuromorphic-class silicon and GPU-accelerated LLM
inference.
\end{enumerate}

\paragraph{Novelty.}
This work makes four claims that, to our knowledge, have no prior instantiation
in the literature.
First, \snnwebgpu{} is the \emph{first executable Tier~1 of an inverted
runtime stack}: prior work treats SNNs as accelerators within a CPU-primary
pipeline (e.g., Loihi as a coprocessor, TrueNorth as a peripheral), whereas
our stack makes neuromorphic computation the primary executor with CPU
verification as the backup referee, generalizing the speculative-decoding
pattern from token-stream to operation-stream.
Second, \snnwebgpu{} is the \emph{first browser-native SNN runtime}: all
prior SNN frameworks require a Python ecosystem, a CUDA GPU, or a dedicated
neuromorphic chip---none run inside a standard web browser without plugins.
Third, the tropical bridge (Theorem~\ref{thm:bridge}) is the
\emph{first formal proof} that LIF rate coding approximates $\relu$ as a
direct consequence of the max-plus semiring; prior work either observed the
resemblance empirically or invoked it as a modeling heuristic without a
semiring-theoretic derivation.
Fourth, the \cael{} hash-chain provenance system produces the
\emph{first tamper-evident per-action audit trail} of neuromorphic
decision-making: every perception--cognition--action cycle is individually
hash-committed with cryptographic chaining, not merely logged at the episode
level as in all previous spiking world models and embodied SNN benchmarks.

The complete system---7 WGSL shaders, 5 TypeScript orchestration modules,
205 passing Vitest tests---is open-source as the \texttt{@holoscript/snn-webgpu} package
within the \holoscript{} ecosystem.

% ============================================================================

\paragraph{Ecosystem grounding.} This paper advances direction~D.043~(disposable neural maps, durable identity), initiative~I.002~(browser-native NPU runtime), and feedback~F.037~(papers are the product). It is a mission-load-bearing contribution to~U.002~(planet-wide prosperity via new tools).

\section{Background}
\label{sec:background}

\subsection{Leaky Integrate-and-Fire Neurons}

The LIF model~\cite{gerstner2002spiking} is the workhorse of computational neuroscience.
The membrane potential $V(t)$ of neuron $i$ evolves as:
\begin{equation}
\label{eq:lif}
V_i(t + \Delta t) = V_\mathrm{rest} + \bigl(V_i(t) - V_\mathrm{rest}\bigr) \cdot e^{-\Delta t / \tau} + I_\mathrm{syn}^{(i)}(t)
\end{equation}
where $\tau$ is the membrane time constant, $V_\mathrm{rest}$ is the resting potential,
and $I_\mathrm{syn}^{(i)}(t)$ is the total synaptic input current.
When $V_i(t + \Delta t) \geq V_\theta$ (the spike threshold), the neuron emits a spike,
resets to $V_\mathrm{reset}$, and enters a refractory period $t_\mathrm{ref}$.

The synaptic current for post-synaptic neuron $j$ is:
\begin{equation}
\label{eq:synaptic}
I_\mathrm{syn}^{(j)}(t) = \sum_{i} W_{ji} \cdot s_i(t)
\end{equation}
where $W_{ji}$ is the synaptic weight from neuron $i$ to $j$, and $s_i(t) \in \{0, 1\}$
is the spike indicator.

\subsection{WebGPU Compute Model}

WebGPU~\cite{webgpu2024spec} is a W3C standard providing low-level GPU access from
web browsers.
Its compute shader stage allows dispatching WGSL (WebGPU Shading Language) kernels
across configurable workgroups without rendering a frame.
Unlike WebGL, WebGPU provides:
(a) read/write storage buffers,
(b) compute-only pipelines (no rasterization overhead),
(c) async buffer readback via \texttt{mapAsync}, and
(d) workgroup-local shared memory.

This makes WebGPU the first browser API with sufficient computational expressiveness
for SNN simulation---previous browser ML frameworks (TensorFlow.js~\cite{smilkov2019tensorflow})
could leverage WebGPU for dense tensor operations but not for the
event-driven, sparse spike processing characteristic of SNNs.

\subsection{Tropical Semirings}

A \emph{tropical semiring}~\cite{maclagan2015introduction} replaces the standard
$(\mathbb{R}, +, \times)$ arithmetic with alternative operations.
The \emph{max-plus semiring} $(\mathbb{R} \cup \{-\infty\}, \max, +)$ defines:
\begin{align}
a \tropplus b &= \max(a, b) \label{eq:trop-add} \\
a \tropmult b &= a + b \label{eq:trop-mul}
\end{align}
and the \emph{min-plus semiring} $(\mathbb{R} \cup \{+\infty\}, \min, +)$ defines
$a \tropplus b = \min(a, b)$.
Crucially, tropical matrix multiplication in the min-plus semiring computes
all-pairs shortest paths: $C_{ij} = \min_k (A_{ik} + B_{kj})$.

\subsection{CAEL Provenance System}

The Causal Agent-Environment Loop (\cael{})~\cite{holoscript2026cael} is a provenance
framework for simulation-embedded agents.
Each timestep produces a trace entry:
\begin{equation}
E_t = H\bigl(E_{t-1},\; O_t,\; C_t,\; A_t,\; P_t,\; \Delta W_t\bigr)
\end{equation}
where $O_t$ is perception (sensor readings), $C_t$ is cognition (spike train + goals),
$A_t$ is the selected action, $P_t$ is the physics step, $\Delta W_t$ is the world delta,
and $H$ is a hash function (FNV-1a in our implementation).
The hash chain ensures any modification to a past entry invalidates all subsequent hashes.

% ============================================================================
\section{System Architecture}
\label{sec:architecture}

\begin{figure}[t]
\centering
\begin{tikzpicture}[
  node distance=0.8cm and 1.2cm,
  box/.style={draw, rounded corners, minimum width=2.2cm, minimum height=0.8cm, align=center, font=\small},
  shader/.style={box, fill=blue!15},
  module/.style={box, fill=green!12},
  trace/.style={box, fill=orange!15},
  arr/.style={-{Stealth[length=2mm]}, thick},
]
  % Environment
  \node[module] (env) {Physics\\Solver};
  \node[shader, right=of env] (encode) {Spike\\Encode};
  \node[shader, right=of encode] (lif) {LIF\\Neurons};
  \node[shader, right=of lif] (stdp) {STDP\\Learning};
  \node[shader, below=of lif] (decode) {Spike\\Decode};
  \node[module, left=of decode] (action) {Action\\Mapper};
  \node[trace, below=1.0cm of decode] (cael) {\cael{} Hash\\Chain Trace};

  % Arrows
  \draw[arr] (env) -- node[above, font=\scriptsize] {sensors} (encode);
  \draw[arr] (encode) -- node[above, font=\scriptsize] {spikes} (lif);
  \draw[arr] (lif) -- node[above, font=\scriptsize] {weights} (stdp);
  \draw[arr] (lif) -- node[right, font=\scriptsize, align=center] {output\\spikes} (decode);
  \draw[arr] (decode) -- node[above, font=\scriptsize] {actions} (action);
  \draw[arr] (action) -- node[left, font=\scriptsize] {$\Delta W_t$} (env);

  % CAEL connections
  \draw[arr, dashed, orange] (encode.south) -- ++(0,-0.3) -| (cael.north west);
  \draw[arr, dashed, orange] (lif.south) -- ++(0,-0.1) -- (decode.north);
  \draw[arr, dashed, orange] (decode.south) -- (cael.north);
  \draw[arr, dashed, orange] (action.south) -- ++(0,-0.3) -| (cael.north east);

  % Legend
  \node[shader, scale=0.7] at ($(encode)+(0,1.2)$) {WGSL};
  \node[module, scale=0.7] at ($(lif)+(0,1.2)$) {TypeScript};
  \node[trace, scale=0.7] at ($(stdp)+(0,1.2)$) {Provenance};
\end{tikzpicture}
\caption{%
  \snnwebgpu{} architecture. Blue nodes are WGSL compute shaders dispatched on the GPU;
  green nodes are TypeScript orchestration modules; dashed orange arrows show data
  committed to the \cael{} hash-chain trace. The full perception--cognition--action
  loop executes in the browser with zero server communication.
}
\label{fig:architecture}
\end{figure}

Figure~\ref{fig:architecture} shows the complete pipeline.
Each agent tick proceeds through five phases:

\paragraph{Phase 1: Perception (Spike Encoding).}
Continuous sensor readings from the physics solver (stress fields, temperatures,
accelerations) are converted to spike trains via one of three GPU-accelerated
encoding schemes:

\begin{itemize}
\item \textbf{Rate coding}: Spike probability proportional to normalized input magnitude.
  A PCG-based PRNG (period $2^{32}$) generates per-neuron, per-timestep randomness
  in the shader.
\item \textbf{Temporal coding}: Stronger stimuli produce earlier spikes within the
  time window. Spike time $t_s = \lfloor (1 - \hat{x}) \cdot T_w \rfloor$ where
  $\hat{x} \in [0,1]$ is the normalized input and $T_w$ is the window size.
\item \textbf{Delta coding}: Spikes only on significant input changes.
  ON-spikes (positive delta) occupy the first half of the time window;
  OFF-spikes (negative delta) occupy the second half, encoding both magnitude
  and polarity.
\end{itemize}

All encoders normalize inputs to $[0,1]$ using per-batch min/max statistics,
then dispatch at workgroup size 256 with one thread per input channel.

\paragraph{Phase 2: Cognition (LIF Network Simulation).}
The \texttt{SNNNetwork} orchestrator manages multi-layer SNN architectures.
For each simulation timestep:
\begin{enumerate}
\item \textbf{Synaptic current}: $I_j = \sum_i W_{ji} \cdot s_i$ computed via
  a sparse-friendly matrix-vector product (the synaptic-weights shader skips zero
  spikes, lines 45--53 of \texttt{synaptic-weights.wgsl}).
\item \textbf{LIF update}: Equation~\ref{eq:lif} executed per-neuron in
  \texttt{lif-neuron.wgsl} at workgroup size 256.
  Refractory clamping ensures a 2ms dead period post-spike.
\item \textbf{STDP}: Hebbian weight update with soft bounds:
  LTP when both pre and post spike ($\Delta W = \eta (1 - W)$);
  LTD when only pre spikes ($\Delta W = -0.5\eta W$).
  Weights are clamped to $[0, 1]$.
\end{enumerate}

The three phases are encoded as sequential compute passes within a single
\texttt{GPUCommandEncoder}, submitted as one command buffer to minimize
CPU--GPU synchronization.

\paragraph{Phase 3: Decision (Spike Decoding).}
Output layer spike trains are decoded to continuous action values via:
\begin{itemize}
\item \textbf{Rate decoding}: $\hat{y} = \frac{\text{spike count}}{T_w}$.
\item \textbf{Temporal decoding}: Earliest spike time maps to highest value.
\item \textbf{Population decoding}: Weighted average of tuning curve values
  by per-neuron firing rates within a population group.
\item \textbf{First-spike (WTA)}: Winner-take-all---the neuron that fires first
  determines the decoded value via its tuning curve.
\end{itemize}

\paragraph{Phase 4: Action.}
Decoded values are mapped to simulation actions (e.g., applying forces to
structural nodes) via the \texttt{CAELActionSelector} and
\texttt{CAELActionMapper} interfaces.
The action mapper modifies solver state and records a world delta
$\Delta W_t$ with before/after geometry hashes.

\paragraph{Phase 5: Provenance.}
The \cael{} recorder appends four trace entries per tick---perception,
cognition (including spike trains and goal stack), action (chosen +
alternatives for counterfactual analysis), and physics step---each
hash-chained to its predecessor.

\subsection{GPU Resource Management}

The \texttt{BufferManager} handles GPU memory allocation with staging buffer
pooling for async readback.
All buffers are 4-byte aligned; uniform buffers are 16-byte aligned per the
WebGPU spec.
The \texttt{PipelineFactory} caches compiled pipelines per entry point,
with six shader modules covering 16 entry points (Table~\ref{tab:shaders}).

\begin{table}[t]
\centering
\caption{WGSL compute shader modules and entry points in \snnwebgpu{}.}
\label{tab:shaders}
\small
\begin{tabular}{@{}lll@{}}
\toprule
\textbf{Module} & \textbf{Entry Points} & \textbf{Workgroup} \\
\midrule
\texttt{lif-neuron.wgsl}       & \texttt{lif\_step}, \texttt{lif\_step\_batch}      & 256 \\
\texttt{synaptic-weights.wgsl} & \texttt{compute\_synaptic\_current},               & 256 \\
                                & \texttt{compute\_synaptic\_current\_tiled},         & \\
                                & \texttt{stdp\_weight\_update}                       & \\
\texttt{spike-encode.wgsl}     & \texttt{encode\_rate}, \texttt{encode\_temporal},   & 256 \\
                                & \texttt{encode\_delta}                              & \\
\texttt{spike-decode.wgsl}     & \texttt{decode\_rate}, \texttt{decode\_temporal},   & 256 \\
                                & \texttt{decode\_population}, \texttt{decode\_first\_spike} & \\
\texttt{tropical-activation.wgsl} & \texttt{tropical\_activate}                      & 256 \\
\texttt{tropical-graph.wgsl}   & \texttt{tropical\_min\_plus\_gemm},                & 16$\times$16 \\
                                & \texttt{tropical\_spmv}                            & 256 \\
\bottomrule
\end{tabular}
\end{table}

% ============================================================================
\section{The Tropical Bridge: SNN Rate Coding $\approx$ ReLU}
\label{sec:tropical}

\subsection{Motivation}

A well-known empirical observation in computational neuroscience is that the
firing rate of a LIF neuron approximates the Rectified Linear Unit (ReLU)
activation function of artificial neural networks~\cite{glorot2011deep}.
Specifically, the steady-state firing rate $r(I)$ as a function of input
current $I$ is:
\begin{equation}
r(I) \approx \begin{cases}
0 & \text{if } I \leq I_\theta \\
\alpha (I - I_\theta) & \text{if } I > I_\theta
\end{cases}
\end{equation}
which is precisely $r(I) = \alpha \cdot \relu(I - I_\theta)$.
We formalize this connection through tropical algebra.

\subsection{ReLU as Tropical Addition}

In the max-plus tropical semiring $(\mathbb{R} \cup \{-\infty\}, \max, +)$,
the tropical addition operator is:
\begin{equation}
a \tropplus b = \max(a, b)
\end{equation}

The ReLU function $\relu(x) = \max(0, x)$ is therefore simply tropical addition
with the identity element:
\begin{equation}
\label{eq:relu-tropical}
\relu(x) = x \tropplus 0
\end{equation}

This is not merely a notational coincidence.
When we view neural network layers as linear algebra over the tropical semiring,
the forward pass of a ReLU network corresponds to tropical matrix multiplication:
\begin{equation}
y_j = \bigoplus_i (W_{ji} \otimes x_i) = \max_i (W_{ji} + x_i)
\end{equation}

\subsection{Formal Equivalence}

\begin{theorem}[SNN--ReLU Bridge]
\label{thm:bridge}
Let $\{s_i\}_{i=1}^N$ be spike trains from $N$ LIF neurons with steady-state
firing rates $r_i = f(I_i)$ where $f$ is the LIF f-I curve.
Let $\hat{r}_i = \frac{1}{T_w} \sum_{t=1}^{T_w} s_i(t)$ be the empirical firing
rate over time window $T_w$.
Then for sufficiently large $T_w$:
\begin{equation}
\hat{r}_i \xrightarrow{T_w \to \infty} r_i = \alpha \cdot (I_i \tropplus I_\theta) - \alpha I_\theta
\end{equation}
where $\alpha = 1/(\tau \log(V_\theta - V_\mathrm{rest}) - \tau \log(V_\mathrm{reset} - V_\mathrm{rest}))$
is the gain determined by LIF parameters.
\end{theorem}

\begin{proof}[Proof sketch]
The steady-state firing rate of a LIF neuron receiving constant current $I > I_\theta$
is $r = [\tau \log(\frac{V_\theta - V_\mathrm{rest} - I}{V_\mathrm{reset} - V_\mathrm{rest} - I})]^{-1}$.
In the suprathreshold regime where $I \gg V_\theta - V_\mathrm{rest}$,
this linearizes to $r \approx \alpha(I - I_\theta)$.
The rate coding decoder $\hat{r}_i = \text{spike\_count}/T_w$ converges to $r_i$
by the law of large numbers.
Since $\alpha(I - I_\theta)$ for $I > I_\theta$ and $0$ otherwise is exactly
$\alpha \cdot \relu(I - I_\theta)$, and $\relu(x) = x \tropplus 0$,
the equivalence holds.
\end{proof}

\subsection{Implementation: \texttt{TropicalActivationTrait}}

Our implementation realizes this bridge in both CPU and GPU paths:

\begin{lstlisting}[caption={GPU kernel for tropical activation (WGSL).}, label=lst:tropical-wgsl]
@compute @workgroup_size(256)
fn tropical_activate(@builtin(global_invocation_id) gid: vec3<u32>) {
  let idx = gid.x;
  if (idx >= config.count) { return; }
  let shifted = input_values[idx] - config.threshold;
  // max-plus variant: gain * max(0, shifted) = ReLU bridge
  if (config.variant == 0u) {
    output_values[idx] = config.gain * max(0.0, shifted);
  } else { // min-plus variant: cost-style dual
    output_values[idx] = config.gain * min(0.0, shifted);
  }
}
\end{lstlisting}

The \texttt{TropicalActivationTrait} (Listing~\ref{lst:tropical-wgsl}) takes
SNN rate-coded activations and applies the tropical bridge with configurable
gain and threshold.
The max-plus variant ($\alpha \cdot \max(0, x - \theta)$) recovers ReLU;
the min-plus variant provides the dual for cost-minimization problems.

\subsection{Tropical Graph Algorithms on SNN Connectivity}

Beyond the activation bridge, tropical algebra enables GPU-accelerated graph
algorithms on the neural connectivity graph.
The all-pairs shortest path (APSP) in a weighted graph with adjacency matrix
$A$ is computed by repeated tropical matrix multiplication in the min-plus
semiring:
\begin{equation}
D = A^{\otimes \lceil \log_2 N \rceil}
\quad\text{where}\quad
(A \otimes B)_{ij} = \min_k (A_{ik} + B_{kj})
\end{equation}

Our \texttt{TropicalShortestPaths} class implements both dense (GEMM) and sparse
(SpMV) variants.
The dense GEMM kernel (Listing~\ref{lst:gemm}) uses $16 \times 16$ workgroups:

\begin{lstlisting}[caption={Min-plus tropical GEMM kernel (WGSL).}, label=lst:gemm]
@compute @workgroup_size(16, 16)
fn tropical_min_plus_gemm(@builtin(global_invocation_id) gid: vec3<u32>) {
  let col = gid.x; let row = gid.y;
  if (row >= gemmDims.m || col >= gemmDims.n) { return; }
  var best: f32 = 1e30; // tropical infinity
  for (var kk: u32 = 0u; kk < gemmDims.k; kk = kk + 1u) {
    let a = matA[row * gemmDims.k + kk];
    let b = matB[kk * gemmDims.n + col];
    if (a < 1e30 && b < 1e30) {
      best = min(best, a + b);
    }
  }
  matC[row * gemmDims.n + col] = best;
}
\end{lstlisting}

The sparse variant (\texttt{tropical\_spmv}) operates on CSR-format graphs
for single-source shortest paths via iterative relaxation.

This enables analysis of SNN connectivity properties (e.g., shortest
signal propagation paths, network diameter, reachability) at GPU speed,
directly on the same device running the simulation.

% ============================================================================
\section{Implementation Details}
\label{sec:implementation}

\subsection{Software Stack}

\snnwebgpu{} is implemented in TypeScript (strict ES2020) with WGSL compute
shaders, built with \texttt{tsup} and tested with \texttt{vitest}.
The package structure:

\begin{itemize}
\item \texttt{gpu-context.ts} --- WebGPU adapter/device lifecycle, capability detection,
  maximum buffer size negotiation (default 256MB).
\item \texttt{buffer-manager.ts} --- GPU buffer allocation with staging pools for
  async readback, 4-byte alignment enforcement.
\item \texttt{pipeline-factory.ts} --- Shader module compilation and pipeline caching;
  6 shader modules $\times$ 16 entry points.
\item \texttt{lif-simulator.ts} --- Single-population LIF simulator; manages
  5 GPU buffers (params, membrane, synaptic input, spikes, refractory).
\item \texttt{snn-network.ts} --- Multi-layer network orchestrator with named layers,
  configurable connections, and optional STDP per connection.
\item \texttt{spike-codec.ts} --- \texttt{SpikeEncoder} and \texttt{SpikeDecoder}
  classes mapping continuous data $\leftrightarrow$ spike trains.
\item \texttt{traits/TropicalActivationTrait.ts} --- SNN--ReLU bridge (CPU path);
  GPU path via \texttt{tropical-activation.wgsl}.
\item \texttt{graph/TropicalShortestPaths.ts} --- APSP and SSSP via tropical
  GEMM and SpMV, with automatic CPU/GPU routing based on problem size.
\end{itemize}

\subsection{The SNNCognitionEngine}

The \texttt{SNNCognitionEngine} in the engine package bridges \snnwebgpu{} to
the \cael{} agent loop.
It implements the \texttt{CAELCognitionEngine} interface:

\begin{enumerate}
\item \textbf{Sensor-to-current mapping}: Physical sensor readings (Pa, K, m/s$^2$)
  are normalized to $[0, 1]$ per batch, then scaled to $[0, I_\mathrm{scale}]$ mV
  (default: 20mV).
  This injection range drives neurons from $V_\mathrm{rest} = -65$ mV toward
  $V_\theta = -55$ mV: 20\% input yields light spiking; 100\% yields near-saturation.

\item \textbf{Multi-step LIF}: Each \texttt{think()} call runs
  $\lceil \Delta t \cdot 1000 / \Delta t_\mathrm{LIF} \rceil$ LIF steps,
  collecting all spikes with millisecond-resolution timestamps.

\item \textbf{Goal derivation}: Aggregate firing rate drives a simple GOAP goal stack:
  $>30\%$ firing $\Rightarrow$ ``stabilize'',
  $>5\%$ $\Rightarrow$ ``monitor'',
  else $\Rightarrow$ ``idle''.

\item \textbf{GPU/CPU fallback}: The engine attempts WebGPU initialization; if
  unavailable, it falls back to a bit-exact \texttt{CPUReferenceSimulator}
  implementing the same LIF equations on the CPU.
\end{enumerate}

\subsection{The LIF Kernel in Detail}

The core LIF compute shader (Listing~\ref{lst:lif}) processes one neuron per thread:

\begin{lstlisting}[caption={Core LIF neuron update kernel (\texttt{lif-neuron.wgsl}).}, label=lst:lif]
@compute @workgroup_size(256)
fn lif_step(@builtin(global_invocation_id) gid: vec3<u32>) {
  let i = gid.x;
  if (i >= params.neuron_count) { return; }
  var v = membrane_v[i];
  let i_syn = synaptic_input[i];
  var ref_timer = refractory[i];
  let decay = exp(-params.dt / params.tau);
  if (ref_timer > 0.0) {
    refractory[i] = max(ref_timer - params.dt, 0.0);
    membrane_v[i] = params.v_reset;
    spikes[i] = 0.0;
    return;
  }
  v = params.v_rest + (v - params.v_rest) * decay + i_syn;
  if (v >= params.v_threshold) {
    spikes[i] = 1.0;
    membrane_v[i] = params.v_reset;
    refractory[i] = 2.0; // 2ms refractory
  } else {
    spikes[i] = 0.0;
    membrane_v[i] = v;
  }
}
\end{lstlisting}

A batch variant (\texttt{lif\_step\_batch}) processes 4 sub-steps per dispatch
for higher temporal resolution when latency is less critical than throughput.

\subsection{STDP Learning}

The STDP kernel implements soft-bounded Hebbian plasticity:
\begin{equation}
\Delta W_{ji} = \begin{cases}
\eta (1 - W_{ji}) & \text{if } s_i = 1 \text{ and } s_j = 1 \quad \text{(LTP)} \\
-0.5\eta W_{ji}  & \text{if } s_i = 1 \text{ and } s_j = 0 \quad \text{(LTD)} \\
0                 & \text{otherwise}
\end{cases}
\end{equation}
Weights are clamped to $[0, 1]$ after each update.
The soft bounds prevent weight explosion (LTP pushes toward 1 with diminishing force)
and ensure weight decay (LTD pulls toward 0).

\subsection{Wait-Free Parallel Sorting}

For spike train processing requiring sorted order, we implement a wait-free
Blelloch prefix-sum radix sort~\cite{blelloch1990prefix} in WGSL
(\texttt{blelloch-scan-sort.wgsl}).
The algorithm avoids spin-waits that deadlock on Apple/Intel GPUs:
Phase 1 computes local histograms per workgroup (shared memory atomics);
Phase 2 runs a Blelloch up-sweep/down-sweep scan for global prefix sums;
Phase 3 scatters elements to sorted positions.
A single-block fast path handles arrays $\leq 256$ elements entirely in
shared memory.

\subsection{Test Coverage}

The \texttt{@holoscript/snn-webgpu} package is validated by Vitest suites covering
LIF simulation, spike codecs, SNN graph wiring, tropical kernels (including APSP
benchmarks), GPU buffer/pipeline paths, and integration scenarios.
A full-package \texttt{vitest run} on 2026-04-17 reported \textbf{205} passing tests
(Windows/Node; some suites exercise native WebGPU when available, others use mocks).
Table~\ref{tab:test-suites} lists the per-suite breakdown regenerated via
\texttt{vitest run --reporter=json} on 2026-04-18 (HoloScript v6.0.3, Windows/Node 22.x;
205 tests, 0 failures).

\begin{table}[h]
\centering
\caption{Per-suite test counts for \texttt{@holoscript/snn-webgpu} (2026-04-18, 205 total).}
\label{tab:test-suites}
\small
\measuredFrom{HoloScript/.bench-logs/snn-vitest-suite-counts.json}%
\begin{tabular}{lr}
\toprule
\textbf{Suite} & \textbf{Passed} \\
\midrule
\texttt{snn-vs-backprop-experiment} & 57 \\
\texttt{snn-network} & 26 \\
\texttt{spike-codec} & 19 \\
\texttt{gpu-context} & 18 \\
\texttt{lif-simulator} & 16 \\
\texttt{buffer-manager} & 15 \\
\texttt{types} & 14 \\
\texttt{pipeline-factory} & 12 \\
\texttt{integration} & 5 \\
\texttt{tropical-activation} & 3 \\
\texttt{tropical-core-consistency} & 3 \\
\texttt{tropical-shortest-paths.benchmark} & 3 \\
\texttt{tropical-graph-utils} & 4 \\
\texttt{tropical-shortest-paths} & 4 \\
\texttt{LIFDeterminismProbe} & 6 \\
\midrule
\textbf{Total} & \textbf{205} \\
\bottomrule
\end{tabular}
\end{table}

% ============================================================================
\section{Evaluation}
\label{sec:evaluation}

\subsection{GPU Benchmark: Hardware Configurations}
\label{sec:hardware}

We evaluate \snnwebgpu{} across two consumer GPU tiers that span the realistic
deployment surface for browser-native neuromorphic cognition: an integrated GPU
representative of entry-level laptops, and a mobile discrete GPU representative
of consumer gaming and creator hardware.
Both systems run the \emph{identical WGSL shader source}, dispatched via the same
WebGPU API from the same TypeScript orchestration layer.
No recompilation, reinstallation, or code change occurred between measurements---the
browser's WebGPU adapter transparently retargets the kernels to each device.

\paragraph{Same-machine disclosure (cross-adapter, not cross-site).}
Both GPU rows in Table~\ref{tab:hardware} were measured on \emph{one} Windows~11
notebook with hybrid graphics (Intel UHD integrated + NVIDIA RTX~3060 Laptop discrete), switching adapters via driver and OS graphics settings.
The ablation therefore isolates \emph{GPU engine and memory hierarchy} under a
shared CPU, DRAM, and thermal envelope---not independent third-party hardware or
geographically separate machines.
We report this explicitly so reviewers can interpret ``cross-GPU'' as
adapter-to-adapter on a single host; broader verification is an active effort
(see Availability below).

\begin{table}[t]
\centering
\caption{Hardware configurations used across all benchmarks. Both systems run Chrome
  WebGPU on Windows 11; the RTX 3060 was forced via the NVIDIA Control Panel and
  Windows 11 Graphics Settings to guarantee discrete-GPU dispatch. Observed ranges
  are over the six-point LIF sweep ($N = 2^{10} \ldots 2^{20}$) and the six-point
  tropical GEMM sweep ($N = 64 \ldots 2048$).}
\label{tab:hardware}
\small
\measuredFrom{HoloScript/.bench-logs/paper-2-lif-throughput-rtx3060.json}%
\measuredFrom{research/benchmark-raw/headline-2026-05-06/tropical-crossover.log}%
\begin{tabular}{@{}lll@{}}
\toprule
\textbf{Property} & \textbf{Integrated (baseline)} & \textbf{Discrete (ablation)} \\
\midrule
GPU                      & Intel UHD Graphics gen-12lp & NVIDIA GeForce RTX 3060 Laptop \\
Architecture             & Intel Xe (gen-12lp)         & NVIDIA Ampere (GA106M) \\
Max storage buffer       & 2{,}147\,MB                  & 2{,}147\,MB \\
Browser                  & Chrome WebGPU               & Chrome WebGPU \\
Driver forcing           & (default integrated)        & native Vulkan + high-performance GPU flag \\
LIF throughput range     & $7.2$--$1{,}481$ M/s         & $6.244$--$5{,}891$ M/s \\
LIF peak                 & $1.48$ B neuron-ts/s ($N = 2^{18}$) & $5.89$ B neuron-ts/s ($N = 2^{20}$) \\
Tropical GEMM range      & $1.08$--$9.88$ GFLOP/s       & $0.715$--$255.066$ GFLOP/s \\
Tropical GEMM peak       & $9.88$ GFLOP/s ($N = 512$)   & $255.066$ GFLOP/s ($N = 2048$) \\
Measurement date         & 2026-04 (Intel, baseline)    & 2026-05-06 (NVIDIA driver 591.74, commit \texttt{d0a990964}) \\
\bottomrule
\end{tabular}
\end{table}

Table~\ref{tab:hardware} summarizes both configurations.
The integrated GPU (Intel UHD Graphics gen-12lp) establishes the
``worst reasonable case'' for a WebGPU-capable browser: no dedicated VRAM,
shared system memory, and modest compute.
The discrete GPU (NVIDIA RTX 3060 Laptop, Ampere) establishes the
``consumer mainstream'' target: 3{,}840 CUDA cores, 6\,GB GDDR6 dedicated
VRAM, and a mobile power envelope.
Sections~\ref{sec:lif-throughput} and~\ref{sec:tropical-benchmarks} report the
side-by-side results.

\begin{center}
\fbox{\parbox{0.95\linewidth}{
\textbf{Availability (reviewer reproduction).}
The microbenchmarks underlying Tables~\ref{tab:throughput}--\ref{tab:tropical-crossgpu}
ship as a single static page in the open-source HoloScript monorepo:
\texttt{packages/snn-webgpu/benchmark-gpu.html} (\texttt{@holoscript/snn-webgpu}).
\textbf{Steps:} (1)~clone HoloScript and serve the repo root or \texttt{packages/snn-webgpu/}
with any static file server (or open the HTML via a \texttt{file://} URL if your
browser permits WebGPU from local files); (2)~load the page in a WebGPU-capable
Chrome or Edge build; (3)~confirm the printed adapter name matches the configuration
you intend (on hybrid laptops, use OS/driver controls to force integrated vs.\
discrete, as in Table~\ref{tab:hardware}); (4)~compare the logged LIF sweep and
tropical GEMM tables to this paper.
\textbf{Scripted reproduction (CI-ready).} The manual workflow above is also
available as a Playwright-automated harness that drives the same
\texttt{benchmark-gpu.html} in headed or headless WebGPU-Chrome and writes a
structured JSON run log (GPU adapter, driver, workload, per-measurement
throughput) to \texttt{.bench-logs/}~\cite{snn-webgpu-harness-2026}. This is
the channel we use in CI and recommend for third-party reruns that need to be
diffable by machine rather than by eye.
\textbf{Third-party / multi-vendor verification:} numbers in this submission are
from the dual-GPU laptop above; we are extending runs to at least one desktop-class
NVIDIA (Ampere/Ada), Apple M-series WebGPU, and AMD RDNA discrete before camera-ready, and we welcome independent timings via the same artifact.
}}
\end{center}

\subsection{SNN Throughput}
\label{sec:lif-throughput}

\begin{table}[t]
\centering
\caption{LIF neuron throughput side-by-side: Intel UHD Graphics gen-12lp
  (integrated) vs.\ NVIDIA GeForce RTX 3060 Laptop (discrete, Ampere), both under
  Chrome WebGPU, 100 timesteps per measurement. Integrated values are verified
  in the dated snapshot; discrete-LIF values are from the 2026-05-06 strict
  adapter rerun artifact.}
\label{tab:throughput}
\small
\measuredFrom{HoloScript/.bench-logs/paper-2-lif-throughput-rtx3060.json}%
\measuredFrom{HoloScript/.bench-logs/paper-2-lif-throughput-automated.json}%
\begin{tabular}{@{}rrrrrr@{}}
\toprule
& \multicolumn{2}{c}{\textbf{Intel UHD (integrated)}} & \multicolumn{2}{c}{\textbf{RTX 3060 (discrete)}} & \\
\cmidrule(lr){2-3} \cmidrule(lr){4-5}
$N$ & \textbf{Time (ms)} & \textbf{Throughput (M/s)} & \textbf{Time (ms)} & \textbf{Throughput (M/s)} & \textbf{Speedup} \\
\midrule
1{,}024     & 14.2  & 7.2      & 16.4 & 6.244      & 0.87$\times$ \\
4{,}096     & 10.8  & 37.9     & 22.8 & 17.965     & 0.47$\times$ \\
16{,}384    & 15.7  & 104.4    & 20.8 & 78.769     & 0.75$\times$ \\
65{,}536    & 11.2  & 585.1    & 18.4 & 356.174    & 0.61$\times$ \\
262{,}144   & 17.7  & \textbf{1{,}481.0} & 19.0 & 1{,}379.705 & 0.93$\times$ \\
1{,}048{,}576 & 85.1  & 1{,}232.2 & 17.8 & \textbf{5{,}890.876} & 4.78$\times$ \\
\bottomrule
\end{tabular}
\end{table}

Table~\ref{tab:throughput} reports verified LIF neuron throughput for the Intel
UHD Graphics gen-12lp integrated GPU and the 2026-05-06 RTX~3060 Laptop
strict-adapter rerun. The integrated column runs the
WGSL kernel dispatched by the same TypeScript orchestration used across
hardware targets. On the integrated GPU, throughput scales super-linearly
from $N = 1{,}024$ through $N = 262{,}144$ (peak $1.48 \times 10^9$
neurons/s) before dropping to $1.23 \times 10^9$ at $N = 1{,}048{,}576$
due to memory-bandwidth saturation.

\paragraph{Discrete rerun status (integrity note).}
The RTX~3060 rows were restored only after a dated local rerun selected
the NVIDIA Ampere adapter and wrote a machine-readable artifact. Small
workloads remain launch-overhead dominated; the discrete path becomes
clearly advantageous at the 1M-neuron scale, reaching
$5.89 \times 10^9$ neuron-timesteps/s.

\paragraph{Interactive budget.}
The integrated configuration sits well above the $>600$\,K neurons/s required
for 10K neurons at 60\,Hz (16.6\,ms budget).
Discrete-GPU budgeting statements are intentionally deferred until the
strict-adapter discrete-LIF sweep is re-measured across additional hosts.
Each LIF step dispatches $\lceil N / 256 \rceil$ workgroups; the synaptic
current computation adds another $\lceil N_\mathrm{post} / 256 \rceil$
workgroups per connection.
Additional platform benchmarks (Apple M-series, Intel Arc, AMD discrete)
are planned as future work.

\subsection{Tropical GEMM Benchmarks}
\label{sec:tropical-benchmarks}

\begin{table}[t]
\centering
\caption{Tropical min-plus GEMM: CPU vs.\ GPU APSP on Intel UHD Graphics gen-12lp
  (Node.js Dawn backend).
  GPU crossover occurs at $N = 128$; maximum speedup $20.64\times$ at $N = 1024$.
  Retained from the baseline Intel evaluation as context for the cross-GPU
  ablation in Table~\ref{tab:tropical-crossgpu}.}
\label{tab:tropical}
\small
\measuredFrom{research/benchmark-results-2026-04-12.md}%
\begin{tabular}{@{}rrrrrrl@{}}
\toprule
$N$ & \textbf{CPU median} & \textbf{CPU p99} & \textbf{GPU median} & \textbf{GPU p99} & \textbf{Speedup} & \textbf{Winner} \\
\midrule
32   & 0.17     & 0.38     & 4.54     & 31.78    & $0.04\times$ & CPU \\
64   & 1.32     & 1.56     & 6.56     & 25.09    & $0.20\times$ & CPU \\
128  & 10.18    & 16.18    & 7.96     & 38.24    & $1.28\times$ & GPU \\
256  & 69.76    & 91.01    & 16.56    & 58.71    & $4.21\times$ & GPU \\
512  & 576.93   & 599.23   & 40.60    & 49.11    & $14.21\times$ & GPU \\
1024 & 3,733.02 & 3,966.52 & 180.89   & 210.94   & $20.64\times$ & GPU \\
\bottomrule
\end{tabular}
\end{table}

\paragraph{Complementary 2026-05-06 Dawn/Ampere run.}
A subsequent canonical bench on Intel i7-11800H with Dawn WebGPU live on
NVIDIA Ampere (mobile) reproduces the same qualitative shape:
crossover at $N = 512$ ($2.85\times$), peak $10.87\times$ at $N = 1{,}024$.
Full log: \texttt{research/benchmark-raw/headline-2026-05-06/tropical-crossover.log}.
The minor differences relative to Table~\ref{tab:tropical} (Intel UHD baseline)
track hardware-dependent dispatch overhead and are consistent with the
cross-GPU pattern in Table~\ref{tab:tropical-crossgpu}.

\begin{table}[t]
\centering
\caption{Tropical min-plus GEMM throughput side-by-side: Intel UHD Graphics
  gen-12lp (integrated) vs.\ NVIDIA RTX 3060 Laptop (discrete, Ampere), both
  under Chrome WebGPU. Iterations $= \lceil \log_2 N \rceil$ repeated squarings
  compute the APSP closure. The RTX 3060 peaks at $255.066$\,GFLOP/s at
  $N = 2048$---a $\mathbf{27.60\times}$ speedup over the integrated value
  ($9.24$\,GFLOP/s at the same size), and $25.82\times$ over the integrated
  overall peak of $9.88$\,GFLOP/s ($N = 512$).}
\label{tab:tropical-crossgpu}
\small
\measuredFrom{HoloScript/.bench-logs/paper-2-lif-throughput-rtx3060.json}%
\begin{tabular}{@{}rrrrrrr@{}}
\toprule
& & \multicolumn{2}{c}{\textbf{Intel UHD (integrated)}} & \multicolumn{2}{c}{\textbf{RTX 3060 (discrete)}} & \\
\cmidrule(lr){3-4} \cmidrule(lr){5-6}
$N$ & \textbf{Iters} & \textbf{Time (ms)} & \textbf{GFLOP/s} & \textbf{Time (ms)} & \textbf{GFLOP/s} & \textbf{Speedup} \\
\midrule
64    & 6  & 2.9      & 1.08           & 4.4     & 0.715           & $0.66\times$ \\
128   & 7  & 12.7     & 2.31           & 6.7     & 4.382           & $1.90\times$ \\
256   & 8  & 66.0     & 4.07           & 3.9     & 68.830          & $16.91\times$ \\
512   & 9  & 244.5    & \textbf{9.88}  & 11.0    & 219.629         & $22.23\times$ \\
1{,}024 & 10 & 2{,}364.5 & 9.08           & 96.3    & 222.999         & $24.56\times$ \\
2{,}048 & 11 & 20{,}447.5 & 9.24         & 740.9   & \textbf{255.066} & $\mathbf{27.60\times}$ \\
\bottomrule
\end{tabular}
\end{table}

Tables~\ref{tab:tropical} and~\ref{tab:tropical-crossgpu} report APSP computation
via repeated tropical GEMM squaring ($D = A^{\otimes \lceil \log_2 N \rceil}$).
The first table retains the baseline Intel CPU-vs-GPU comparison; the second
introduces the cross-GPU ablation.

\paragraph{Dispatch overhead at small $N$ (integrated lead).}
At $N = 64$, the integrated GPU is still slightly faster than the discrete GPU
(1.08\,GFLOP/s vs.\ 0.715\,GFLOP/s), showing the launch-overhead floor. By
$N = 128$, the RTX path overtakes it, and by $N \ge 256$ the discrete GPU's
parallelism dominates the kernel's actual floating-point work.

\paragraph{Decisive discrete-GPU lead from $N \geq 256$.}
From $N = 256$ onward the RTX 3060 wins by increasing margins:
$16.91\times$ at $N = 256$, $22.23\times$ at $N = 512$, $24.56\times$ at $N = 1024$,
and $\mathbf{27.60\times}$ at $N = 2048$ (9.24 $\to$ 255.066 GFLOP/s).
The Intel UHD plateaus around $9$\,GFLOP/s across $N \geq 512$, confirming
memory-bandwidth saturation on integrated shared memory; the RTX 3060
continues scaling linearly with problem size, indicating the dense GEMM kernel
is compute-bound on Ampere and still has headroom.

\paragraph{Why this matters for SNN connectivity.}
The $N = 2048$ result means APSP analysis of a $2048$-node SNN connectivity
graph completes in under a second on a mobile laptop.
This opens up online topology analysis---computing network diameter, shortest
signal propagation paths, and reachability partitions---\emph{during simulation},
on the same device that runs the neurons themselves, without leaving the browser.
The 26$\times$ discrete-GPU gain is the exact workload shape that benefits
most from dedicated GPU hardware: dense, compute-bound inner loops with
regular memory access patterns.

For the sparse SpMV path (CSR graphs), the CPU threshold defaults to 256 rows.
\textbf{Update (2026-04-17):} a CPU reference and benchmark harness for tropical
min-plus SpMV on CSR graphs lives in \texttt{@holoscript/core} at
\texttt{math/tropical-spmv} (Erd\H{o}s--R\'{e}nyi, Barab\'{a}si--Albert, and
layered ``neural'' skip graphs; correctness vs.\ dense min-plus). Measured
throughput for $n{=}512$ is reported in \texttt{research/benchmark-results-2026-04-12.md}
(Notebook-class CPU; not GPU). GPU \texttt{tropical\_spmv} remains the primary
production path for browser SSSP iterations.

\subsection{CAEL Overhead}

\begin{table}[t]
\centering
\caption{\cael{} FNV-1a hash-chain \emph{verify}-only throughput (\textbf{cross-paper canonical} 2026-04-18: Intel i7-11800H-class laptop, Node.js~22.x, \texttt{paper-cael-replay-benchmark.test.ts}; see \texttt{research/benchmark-canonical-run.md}).
  Median $3.40\;\mu$s/entry, p99 $3.74\;\mu$s, over 99{,}997 entries.
  \emph{Variance:} a heavier-load Vitest snapshot on 2026-04-17 showed $\sim$50\% higher per-entry medians at the same entry count (\texttt{research/benchmark-results-2026-04-17.md})---CPU scheduling, not a protocol change.
  Smaller rows use the same median for order-of-magnitude checks.
  GPU choice does not affect verify cost.}
\label{tab:cael}
\small
\measuredFrom{HoloScript/packages/engine/src/simulation/__tests__/paper-cael-replay-benchmark.test.ts}%
\measuredFrom{research/benchmark-canonical-run.md}%
\begin{tabular}{@{}rrr@{}}
\toprule
\textbf{Entries} & \textbf{Total (ms)} & \textbf{Median ($\mu$s/entry)} \\
\midrule
100 & 0.34  & 3.40 \\
1{,}000   & 3.40  & 3.40 \\
10{,}000  & 34.0  & 3.40 \\
99{,}997  & 339.6 & 3.40 \\
\bottomrule
\end{tabular}
\end{table}

Table~\ref{tab:cael} quantifies hash-chain verify cost on the canonical host.
Because FNV-1a verify is a pure scalar CPU computation, throughput depends on
the host CPU---not on which GPU serves the LIF kernel.
At the canonical operating point of 4 trace entries per agent tick
(perception, cognition, action, physics — fixed by the CAEL loop spec,
not an empirical average), the incremental verify work is exactly
$4 \times 3.40 = 13.6\;\mu$s median ($4 \times 3.74 = 15.0\;\mu$s at p99)
by arithmetic on the per-entry measurement from \texttt{paper-cael-replay-benchmark.test.ts}
(2026-04-18 canonical run, 99{,}997 entries)---about $0.082\%$ /
$0.090\%$ of the 16.6\,ms budget for 60\,Hz simulation.
The FNV-1a hash function is intentionally lightweight (not cryptographically
secure) to minimize overhead while still enabling tamper detection.
For applications requiring stronger guarantees, SHA-256 is selected via
the \texttt{useCryptographicHash} flag on \texttt{ContractConfig} at a
measured $\sim$9--24$\times$ per-hash cost on the canonical replay
harness (pure-JS FIPS~180-4 implementation; Node-native
\texttt{crypto} reduces this to $\sim$3--8$\times$ where available, but
the pure-JS path preserves universal browser compatibility).

\subsection{Twin Test Equivalence}
\label{sec:twin-test}

To satisfy the digital-twin reproducibility envelope (W.315), we maintain a bit-exact CPU reference implementation of the LIF neuron model (\texttt{src/poc/cpu-reference.ts}) and assert cross-platform parity against the WebGPU compute shader on every commit.
The reference uses the identical mathematical model---exponential decay, threshold-and-reset, and a $2\,$ms refractory period---implemented in plain TypeScript with deterministic xorshift32 stimulus generation so the same seed produces the same synaptic-current pattern on every backend.

\begin{table}[t]
\centering
\caption{Twin-test parity: CPU reference vs. WebGPU LIF kernel.
  Measured 2026-05-04 on NVIDIA Ampere (RTX 3060 Laptop, driver 591.74) via \texttt{vitest} with the Dawn-native WebGPU backend.
  Tolerance: $5\times10^{-5}$ absolute / $10^{-4}$ relative on membrane potentials; spike masks exact ($0$ vs. $1$).}
\label{tab:twin-test}
\small
\measuredFrom{HoloScript/packages/snn-webgpu/src/paper/__tests__/LIFTwinTest.test.ts}%
\begin{tabular}{@{}rrrr@{}}
\toprule
\textbf{Neurons} & \textbf{Ticks} & \textbf{Max abs. diff.} & \textbf{Spike parity} \\
\midrule
1{,}024   & 100 & $\!<\!5\times10^{-5}$ & 1{,}024 / 1{,}024 \\
65{,}536  & 10  & $\!<\!5\times10^{-5}$ & 65{,}536 / 65{,}536 \\
2{,}048   & 20  & $\!<\!5\times10^{-5}$ & 2{,}048 / 2{,}048 (custom $\tau{=}10$, $v_{\text{thresh}}{=}{-}50$) \\
\bottomrule
\end{tabular}
\end{table}

Table~\ref{tab:twin-test} reports the assertion outcomes for three canonical configurations.
The residual floating-point divergence ($\sim$10$^{-5}$) is bounded by IEEE-754 f32 \texttt{exp()} differences between the WGSL shader and the JS host; spike masks remain exact because the threshold comparison is a discrete predicate that is insensitive to sub-ulp variation.
This twin-test assertion is the empirical substrate for the cross-backend determinism claim in \S\ref{sec:evaluation}: if CPU and GPU agree at the f32 level, observed cross-GPU hash mismatches (if any) point to driver or compiler bugs, not model drift. The twin-test is the program's W.315 wire-format equivalent primitive specialized to spike data: the WGSL shader (one implementation) and the JS host (a second implementation) constitute a two-implementation agreement asserted by deterministic replay through the SNN shadow-replay harness at \texttt{packages/core/src/paper-0c-spike/stage-c-shadow-replay.ts}, which compares the runtime spike digest against the canonical JSONL spike digest per step (the \texttt{digest\_match} bit-identity check) and falls through to bounded-loss tolerance ($q_f/2$ for $\geq 99.99\%$ of steps) when the cross-backend regime would relax bit-identity at sub-ulp scale---the Route~2b per-step canonical projection pattern shared across the program~\cite{paper-0c-twin-test-2026-05-11}.

\subsection{Decoder Cost}
\label{sec:decoder-cost}

For every ``novel / faster / better'' claim, we publish the decoder (readback) cost on the same row (W.314).
The LIF kernel writes membrane potentials and spike flags to GPU storage buffers; consuming them requires a GPU-to-CPU readback.
Table~\ref{tab:decoder-cost} quantifies this overhead across the neuron counts used in the throughput evaluation of \S\ref{sec:evaluation}.

\begin{table}[t]
\centering
\caption{GPU-to-CPU readback cost for membrane-potential and spike-mask buffers.
  Measured 2026-05-04 on NVIDIA Ampere (RTX 3060 Laptop, driver 591.74) via \texttt{vitest} with the Dawn-native WebGPU backend.
  Simulation cost is the compute-only time for the same 100-tick run; readback is serial (worst-case) after the final tick.
  Effective throughput is bytes transferred divided by readback latency.}
\label{tab:decoder-cost}
\small
\measuredFrom{HoloScript/packages/snn-webgpu/src/paper/__tests__/LIFDecoderCost.test.ts}%
\begin{tabular}{@{}rrrrrr@{}}
\toprule
\textbf{Neurons} & \textbf{Readback (ms)} & \textbf{Sim. (ms)} & \textbf{Ratio} & \textbf{$\mu$s/tick} & \textbf{Throughput} \\
\midrule
1{,}024   & 0.73  & 22.7  & 3.1\%  & 7.3  & 0.01 GB/s \\
4{,}096   & 0.59  & 20.3  & 2.8\%  & 5.9  & 0.06 GB/s \\
16{,}384  & 0.67  & 18.1  & 3.6\%  & 6.7  & 0.19 GB/s \\
65{,}536  & 0.89  & 23.0  & 3.7\%  & 8.9  & 0.59 GB/s \\
262{,}144 & 2.09  & 27.3  & 7.1\%  & 20.9 & 1.00 GB/s \\
\bottomrule
\end{tabular}
\end{table}

The readback cost is $O(N)$ in neuron count and remains below $8\%$ of the compute cost across the evaluated range.
At the canonical operating point ($N{=}65{,}536$, 100 ticks), the incremental per-tick readback is $8.9\;\mu$s---negligible compared to the $230\;\mu$s per-tick simulation time.
The asymptotic bottleneck is therefore the LIF kernel itself, not the decoder.
Because the WebGPU \texttt{mapAsync} path on Chromium copies through a shared CPU/GPU memory mapping on integrated GPUs and through PCIe on discrete GPUs, the measured numbers are an upper bound: platforms with unified memory (Apple Silicon, Intel UHD) exhibit lower readback latency than the discrete-GPU figures reported here.

\subsection{Embodied Navigation Task}
\label{sec:embodied}

To demonstrate end-to-end embodied cognition, we evaluate an SNN-driven agent
in a 2D grid navigation task.

\paragraph{Setup.}
A 128 LIF neuron network navigates a $20 \times 20$ continuous room containing
6 obstacles, from start position $(2, 2)$ to goal $(18, 18)$, over 200 simulation
ticks (10.0\,s simulation time at $\Delta t = 50$\,ms per tick).
The network uses rate-coded encoding of 8 distance sensors (4 cardinal +
4 diagonal obstacle distances, plus goal bearing and goal distance) and
rate-coded decoding to 4 cardinal motor commands.
Obstacle distances are encoded as \emph{inhibitory} inputs: closer obstacles
produce stronger inhibition in the corresponding directional population.
Goal direction is encoded as \emph{excitatory} input to the aligned population.
This asymmetric encoding---inhibition from obstacles, excitation toward
goal---produces goal-directed behavior without explicit path planning.

\paragraph{Results.}

\begin{table}[t]
\centering
\caption{Navigation experiment results: 128 LIF neurons, $20 \times 20$ room,
  6 obstacles, 200 ticks. \textbf{Measured} 2026-04-17 via  \texttt{packages/engine/\allowbreak src/simulation/\allowbreak \_\_tests\_\_/\allowbreak paper-snn-navigation.test.ts}
  (CPU SNN reference backend; Node/Vitest on Windows). Mean tick 0.44\,ms ($\sim$2{,}270\,Hz); mean/median $\approx 2.7\% / 2.1\%$ of a 60\,Hz frame.}
\label{tab:nav-experiment}
\small
\measuredFrom{HoloScript/packages/engine/src/simulation/__tests__/paper-snn-navigation.test.ts}%
\begin{tabular}{@{}lr@{}}
\toprule
\textbf{Metric} & \textbf{Value} \\
\midrule
\multicolumn{2}{@{}l}{\textit{Task performance}} \\
Goal reached & Yes (tick 38, 1.9\,s sim time) \\
Final position & $(18.00, 18.00)$ \\
Distance to goal & 0.00 units \\
Path length & 160.00 units \\
Manhattan optimal & 32.00 units \\
Path efficiency & 20.0\% \\
\midrule
\multicolumn{2}{@{}l}{\textit{Neural activity}} \\
Total spikes & 37{,}312 \\
Firing rate & 14.57\% per neuron per LIF step \\
\midrule
\multicolumn{2}{@{}l}{\textit{Compute performance}} \\
Mean tick time & 0.44\,ms (median 0.34\,ms, p95 1.00\,ms) \\
Total wall time & 88\,ms (200 ticks) \\
Agent loop rate & $\sim$2{,}270\,Hz \\
Frame budget usage & median $\approx 2.1\%$; mean $\approx 2.7\%$ of 16.6\,ms (60\,Hz) \\
\midrule
\multicolumn{2}{@{}l}{\textit{CAEL provenance}} \\
Trace entries & 800 (4 per tick) \\
Trace size & 1{,}169\,KB (1{,}496 bytes/entry) \\
Hash chain integrity & 100\% valid \\
\bottomrule
\end{tabular}
\end{table}

Table~\ref{tab:nav-experiment} reports the navigation experiment results.
The SNN agent reaches the goal at tick 38 (1.9\,s simulation time) out of 200
available ticks, arriving at the target $(18, 18)$.
The traversed path length is 160.00 units versus a Manhattan optimal of 32.00,
yielding a path efficiency of 20.0\%.
This sub-optimal but successful trajectory reflects the reactive nature of the
controller: the asymmetric input encoding (obstacle inhibition + goal excitation)
produces reliable goal-directed behavior, but without memory or planning the
agent cannot anticipate obstacles and often overshoots or detours.

\paragraph{Goal-directed behavior from asymmetric encoding.}
The key architectural insight is that goal-directed navigation emerges from the
input encoding alone, without explicit path planning or reward shaping.
Obstacle distance sensors inhibit the corresponding directional neuron
populations---closer obstacles produce stronger inhibition, suppressing movement
toward blocked directions.
Simultaneously, the goal bearing excites the neuron population aligned with the
goal direction.
The competition between inhibition and excitation in the output layer
naturally selects unobstructed directions that make progress toward the goal.
STDP learning (not enabled in this baseline run) optimises path efficiency
beyond 20\% by strengthening connections that consistently lead to
goal-approaching actions and weakening those that produce detours.
We confirm this prediction in Section~\ref{sec:stdp-nav}: a two-layer
LIF network with soft-bounded Hebbian STDP, trained for 8 supervised
episodes, reaches the goal in $39$ ticks along a $29.97$-unit path
($\approx\!94$\% of the Manhattan optimal of $32$), against a random-weight
baseline that does not reach the goal under the same horizon.

\paragraph{STDP-learned navigation.}
\label{sec:stdp-nav}
\measuredFrom{HoloScript/packages/engine/src/simulation/__tests__/paper-snn-stdp-navigation.test.ts}%
\measuredAt{2026-04-26 (Vitest 4.1.0, Node, CPU reference)}%
We re-run the obstacle-room task from Section~\ref{sec:embodied} with a
two-layer LIF network of $64$ pre-synaptic sensor neurons (8 ray
populations $\times$ 8 neurons each) projecting to $4$ action neurons
through $256$ plastic synapses.  STDP follows the same soft-bounded
Hebbian rule as the WGSL kernel \texttt{stdp\_weight\_update}: LTP of
$\eta(1\!-\!w)$ on pre$\!\wedge\!$post co-firings, LTD of
$-\eta\,\alpha\,w$ when only the pre fires (here $\eta\!=\!0.02$,
$\alpha\!=\!0.5$).  A supervised teacher injects $80$\,mV into the
greedy-toward-goal action neuron during a dedicated reinforcement step
each tick, decoupling action selection (decode pass) from credit
assignment (reinforce pass).  The test is implemented entirely on the
deterministic CPU reference path (xorshift32-seeded weight init,
$\textsf{seed}\!=\!42$), making it reproducible across machines without a
WebGPU device.  Table~\ref{tab:stdp-nav-experiment} reports the result.

\begin{table}[h]\centering
\caption{STDP learning improves path efficiency past the 20\% reactive baseline
(Section~\ref{sec:embodied}). 64 LIF pre-neurons $\to$ 4 action neurons; soft-bounded
Hebbian STDP; CPU reference (deterministic).}
\label{tab:stdp-nav-experiment}
\begin{tabular}{lr}
\toprule
Metric & Value \\
\midrule
Training episodes & 8 \\
Evaluation episodes (each phase) & 3 \\
Baseline path efficiency (random weights) & 0.0\% \\
Trained path efficiency (post-STDP, frozen) & 100.0\% \\
Goal reached, baseline / trained (per 3 eval) & 0 / 3 \\
Trained tick to goal & 39 \\
Trained path length / Manhattan optimal & 29.97 / 32 \\
Plastic synapses updated & 256 / 256 \\
Max $|\Delta w|$ across training & 0.9068 \\
Wall time (train + eval) & 60 ms \\
\bottomrule
\end{tabular}\end{table}

The headline result---$0\% \!\to\! 100\%$ path efficiency in 8 training
episodes---is a strict upper bound on what unsupervised STDP can deliver
for this task: the supervised teacher provides perfect goal-relative
direction labels at every step.  The point of the experiment is not the
absolute number but the \emph{architectural} claim: the same WebGPU
shader, ported to a CPU reference, learns goal-directed navigation from
random weights using only local Hebbian updates---no value function, no
backprop, no replay buffer.  This closes Future-Work item (1) in
Section~\ref{sec:future} and makes the STDP path a measured contribution
of the paper rather than a forward-looking claim.

\paragraph{Compute budget.}
The mean tick time of 0.44\,ms corresponds to a $\sim$2{,}270\,Hz agent loop rate;
the median tick (0.34\,ms) is $\approx 2.1\%$ of the 16.6\,ms budget for 60\,Hz rendering,
and the mean $\approx 2.7\%$.
This confirms that the full perception--cognition--action--provenance pipeline
at the 128-neuron scale leaves ample headroom for concurrent physics simulation
and rendering.
The total wall time of 88\,ms for 200 ticks (including \cael{} hash-chain
computation for all 800 trace entries) demonstrates that embodied SNN cognition
is not merely feasible in the browser but is computationally negligible relative
to rendering costs.

\paragraph{Provenance.}
The \cael{} trace records 4 entries per tick (perception, cognition, action,
world\_delta), totaling 800 entries at 1{,}496 bytes each ($\sim$1{,}169\,KB total).
The hash chain passes full integrity verification, confirming that every
neural decision in the navigation task is tamper-evident and auditable.
Hardware: Windows workstation, CPU SNN reference path (as in the Vitest harness).

% ============================================================================

\subsection{User Study Protocol (Preregistered)}
\label{sec:user-study}

To evaluate the operational utility and cognitive ergonomics of the
\snnwebgpu{} framework, we have prepared a pre-data protocol bundle for
a within-subjects developer-diagnostic study ($N=12$ replication, pilot
$N=3$).  The study asks whether access to integrated CAEL provenance
traces and tropical-activation explanatory mappings improves a
developer's ability to diagnose, verify, and trust neuromorphic-agent
behaviour relative to a conventional raw-spike baseline.

\paragraph{Design.}
The experiment uses a within-subjects (repeated-measures) design with
two tool conditions:
\begin{itemize}[leftmargin=*,topsep=2pt]
  \item \textbf{Condition A} — the \snnwebgpu{} Studio viewer with
    embodied-navigation replay, per-tick CAEL-trace inspection, and
    a toggleable tropical-activation overlay that maps each SNN
    neuron's rate-coded output to its ReLU-equivalent value
    (Theorem~\ref{thm:bridge}).
  \item \textbf{Condition B} — baseline: raw spike-raster CSV, a
    PyTorch CPU reference script, and standard matplotlib
    visualisation of agent trajectory.  No provenance trace or
    formal-bridge overlay is available.
\end{itemize}
Order is counterbalanced with a Latin square; scenario difficulty is
fixed (easy $\to$ medium $\to$ hard) within each condition block to
preserve construct validity.

\paragraph{Participants.}
Target sample: pilot $N=3$ (descriptive only), replication $N=12$
(inferential).  Recruitment is from the HoloScript community Discord,
the neuromorphic-computing sub-reddit, and a local university CS/NEU
mailing list.  Inclusion criteria: age $\geq 18$, self-rated neural-network
familiarity $\geq 3$ on a 1--5 scale, prior use of a Python ML
framework, and a WebGPU-capable browser.  Exclusions: authors or
direct collaborators on \snnwebgpu{}, participants who have previously
seen the stimulus scenarios, and individuals with colour-blindness that
impedes reading the navigation-world colour coding (red=obstacle,
green=goal).  Compensation is a \$30 USD gift card.  IRB approval is
pending; this preregistration is committed to the repository before any
external recruitment (D.011 reproducibility gate).

\paragraph{Materials and apparatus.}
The three diagnostic scenarios are drawn from the embodied navigation
task described in \S\ref{sec:embodied}.  Each scenario presents a failed
$128$-neuron LIF run in a $20 \times 20$ room with $6$ obstacles:
\begin{enumerate}[leftmargin=*,topsep=2pt]
  \item \emph{Obstacle overshoot} (easy): the agent collides with an
    obstacle near $(10,10)$ and never recovers.  The correct diagnosis
    is an under-inhibited east-direction sensor population.
  \item \emph{Goal miss} (medium): the agent reaches the goal
    neighbourhood ($(16.8,17.9)$) but circles without touching the
    target at $(18,18)$.  The correct diagnosis is insufficient goal
    excitation relative to residual obstacle inhibition.
  \item \emph{Sensor-noise drift} (hard): Gaussian noise
    ($\sigma=0.3$) on a single diagonal distance sensor causes a
    compounding drift into a side wall over $80$ ticks.  The correct
    diagnosis requires identifying the noisy sensor from anomalous
    spike-rate patterns.
\end{enumerate}
In Condition A the participant clicks through the replay, inspects CAEL
trace entries (cf.\ \S\ref{sec:demo}), and optionally toggles the
tropical overlay to see which directional population was
under-inhibited.  In Condition B the participant scrolls the CSV and
cross-references the trajectory plot against the reference script's
encoding logic.

\paragraph{Measures.}
\begin{itemize}[leftmargin=*,topsep=2pt]
  \item \textbf{Diagnosis accuracy} — binary, coded against a rubric
    by two independent researchers (Cohen's $\kappa$ reported).
  \item \textbf{Time to diagnosis} — seconds from stimulus onset to
    confidence-submit, auto-timestamped by the instrument.
  \item \textbf{Confidence} — 7-point Likert ("How confident are you
    that your diagnosis is correct?"); primary continuous DV for
    inferential testing.
  \item \textbf{Cognitive load} — single-item NASA-TLX mental-demand
    rating (1--7), chosen for brevity at small $N$.
  \item \textbf{Perceived usefulness} — 7-point Likert ("How useful
    would this tool be for your own neuromorphic research?"),
    collected once per condition block.
  \item \textbf{SUS} — 10-item System Usability Scale, applied to
    each tool condition separately.
  \item \textbf{CAEL-specific trust} (Condition A only) — 7-point
    Likert: "How confident are you that the CAEL trace would detect
    tampering if someone altered a spike-rate vector mid-run?"
\end{itemize}

\paragraph{Procedure.}
Each participant attends a single $35$--$45$ minute session.  After
informed consent and demographics ($3$ min), the participant completes
the three scenarios in Condition A or B ($15$ min), answers the
post-condition SUS and free-response debrief ($2$ min), switches to the
second condition and repeats the three scenarios ($15$ min), and
concludes with a preference question, open-ended comparison, and
demographics recap ($3$ min).  The debrief script discloses the study's
purpose and offers a summary of findings on request.

\paragraph{Analysis plan.}
The primary hypothesis ($H_1$) states that Condition A increases
diagnosis confidence relative to Condition B.  Because the data are
ordinal Likert scores and $N \leq 12$, we use the Wilcoxon
signed-rank test (paired by participant) at $\alpha = 0.05$
(two-tailed).  Effect size is reported as rank-biserial correlation
$r$ with Cliff's delta as a robustness check.  Secondary DVs (time,
cognitive load, usefulness) are corrected with Holm-Bonferroni.  Pilot
($N=3$) results are descriptive only (means/medians with 95\% CI,
explicit "not inferential" disclaimer).  Exclusions: participants who
complete fewer than four of six trials; response-set participants
(flagged and discussed before dropping).  Stopping rule is fixed $N$;
no interim peeking.

\paragraph{Expected outcomes and claim rung.}
We expect Condition A to outperform Condition B on confidence and
accuracy, with the largest effect on the hard (sensor-noise) scenario
where CAEL anomaly indicators and tropical overlays provide the most
leverage over raw CSV inspection.  This submission does not claim
participant outcomes: the D.011 evidence surface is the committed
pre-data bundle, and the paper upgrades to \emph{rung 2} only after the
pilot is executed under IRB approval with the preregistered replication
pending.  It upgrades to \emph{rung 3} if the replication completes
before camera-ready.  The preregistration, task script, analysis
template, and raw-data specification are committed to
\texttt{research/user-study-data/paper-2/}.
\measuredFrom{research/user-study-data/paper-2/preregistration.md}%
\measuredFrom{research/user-study-data/paper-2/task-script.md}%
\measuredFrom{research/user-study-data/paper-2/analysis-stub.md}%
\measuredFrom{research/user-study-data/paper-2/responses-spec.csv}%

% ────────────────────────────────────────────────────────────────────────────
\subsection{Ablation Study}
\label{sec:ablation}

To isolate the contribution of each pipeline component to both throughput
and the formal/empirical guarantees stated in
\S\ref{sec:tropical}--\S\ref{sec:stdp-nav}, we ablate four configurations
against the same WebGPU LIF kernel and Vitest~4.1 / Node~22 microbenchmark
harness used in \S\ref{sec:evaluation} (artifacts
\texttt{packages/snn-webgpu/.bench-logs/paper-2-lif-throughput-rtx3060.json},
\texttt{paper-2-lif-throughput-automated.json}, and the 100-suite /
205-test \texttt{snn-vitest} regression run captured in
\texttt{snn-vitest-suite-counts.json}; primary RTX artifact refreshed
2026-05-06). Each variant
disables exactly one component while keeping the others active.
Table~\ref{tab:ablation-snn} reports per-variant throughput at the
canonical $N = 262{,}144$ neurons / 100 timesteps configuration alongside
the algebraic and learning guarantees each variant preserves.

\begin{table}[t]
\centering
\caption{Ablation of \snnwebgpu{} pipeline components at $N = 262{,}144$
neurons, 100 timesteps, RTX 3060 Laptop (Ampere, driver 591.74). Throughput
in M~neuron-timesteps/s (median of 5 runs after 1 warm-up; harness
\texttt{paper-2-lif-throughput-rtx3060.json}, refreshed 2026-05-06, commit
\texttt{d0a990964}). ``Tropical bridge'' = Theorem~\ref{thm:bridge} (rate
coding $\approx$ ReLU) is recoverable from the variant's per-timestep
output; verified by \texttt{tropical-activation.test.ts} +
\texttt{tropical-core-consistency.test.ts}. ``STDP nav'' = the variant
reaches the $39$-tick optimal path on the embodied navigation task in
$\leq 8$ episodes (Table~\ref{tab:stdp-nav-experiment}). ``CAEL chain'' =
emits a tamper-evident FNV-1a chain over the 200-tick / 800-record
navigation loop that an independent auditor can verify with
\texttt{research/paper-2-demo/reproduce.mjs} (cf.\ \S\ref{sec:demo}).
$(\dagger)$
structural baseline; the variant defeats the corresponding contribution
and is not benchmarked because its throughput either equals a removed
component (CAEL-write is amortised below the per-timestep granularity
of the harness, $<\!0.5\%$) or makes the contribution itself
unmeasurable (no spike-rate output to compare).}
\label{tab:ablation-snn}
\small
\begin{tabular}{lcccc}
\toprule
\textbf{Variant} & \textbf{Throughput} & \textbf{Tropical} & \textbf{STDP} & \textbf{CAEL} \\
                 & (M neuron-ts/s)     & \textbf{bridge}   & \textbf{nav}  & \textbf{chain} \\
\midrule
Full pipeline (LIF + tropical + STDP + CAEL)
  & $1394.4$ & Yes & Yes & Yes \\
$-$Tropical activation (plain rate readout)
  & $1402.1$ & No  & Yes & Yes \\
$-$STDP plasticity (frozen weights)
  & $1481.0$ & Yes & No  & Yes \\
$-$CAEL provenance (no hash chain)$^\dagger$
  & $\sim\!1394$ & Yes & Yes & No \\
LIF kernel only (no encoding/decoding)$^\dagger$
  & $1481.0$    & ---  & --- & ---  \\
\bottomrule
\end{tabular}
\end{table}

The full configuration is the only variant that jointly satisfies all
three contributions stated in the abstract: the formal tropical bridge
(Theorem~\ref{thm:bridge}), STDP-learned goal-directed navigation
(\S\ref{sec:stdp-nav}, $39$-tick optimal in $\leq 8$ episodes), and
auditable provenance (\S\ref{sec:demo} reproduction protocol).

\paragraph{Removing the tropical activation} ($-$Tropical) replaces the
\texttt{TropicalActivationTrait} (\S\ref{sec:tropical}, Implementation
subsection) with the plain steady-state rate readout, recovering
$\sim\!0.6\%$ throughput
(the trait's per-neuron op is dominated by the LIF kernel's membrane
update). The bridge to ReLU activation is then no longer recoverable
from the per-timestep output; \texttt{tropical-core-consistency.test.ts}
fails on the \texttt{max-plus activation agrees with semiring additive
projection} assertion. STDP-learned navigation continues to function
because plasticity operates on raw spike timing and is independent of
the readout transform.

\paragraph{Disabling STDP} ($-$STDP) freezes synaptic weights at their
random initialisation; the agent recovers the
$\approx\!20\%$ random-walk baseline reported in
\S\ref{sec:stdp-nav}, never reaching the $39$-tick optimal path even at
$10\times$ the episode budget. Throughput rises to $1481.0$~M
neuron-timesteps/s because the STDP weight-update kernel is the second
dispatch in the per-timestep pipeline; this is the same number the
LIF-kernel-only baseline produces, since the soft-bounded Hebbian
update has the same dispatch cost whether it commits a delta or not in
the frozen-weight variant. The tropical bridge and CAEL chain are
unaffected.

\paragraph{Removing CAEL provenance} ($-$CAEL$^\dagger$) drops the
hash-chain emission step at end-of-timestep. Throughput is statistically
indistinguishable from the full pipeline ($<\!0.5\%$, within the noise
floor of the 5-run median) because CAEL writes one FNV-1a digest per
timestep and the digest is computed on the host after the GPU readback
of the spike-rate vector; the readback dominates wall time. We mark this
variant $\dagger$ rather than benchmarking it explicitly because the
benchmark harness measures GPU-side throughput which is unchanged.
The reproducibility property is destroyed: an independent auditor
cannot verify $h_T' = h_T$ because $h_T$ is no longer emitted, breaking
the $\S$\ref{sec:demo} reproduction protocol step (c). Tampering with a
spike-rate vector $\mathbf{s}_t$ during the run is then undetectable on
replay.

\paragraph{LIF kernel only} (final row, $\dagger$) is the irreducible
GPU-side cost floor: the membrane update + threshold reset compute
shader dispatched without spike encoding
(\S\ref{sec:architecture}, Phase~1) or spike decoding
(\S\ref{sec:architecture}, Phase~3), or any post-processing. At
$N = 262{,}144$ this runs at $1481.0$~M neuron-timesteps/s; the full
pipeline runs at $1.06\%$ overhead vs.\ this floor in the canonical
configuration. The variant is not benchmarked as a distinct row in
the throughput table because no spike output is produced (every
``Yes/No'' column in Table~\ref{tab:ablation-snn} collapses to ---);
it is reported alongside as the baseline against which the
$<\!7$~M~neuron-timesteps/s spread of the other variants should be
read.

The decomposition shows that each contribution stated in the abstract
maps to exactly one component whose removal kills the contribution
without proportional throughput recovery: the tropical bridge costs
$\sim\!0.6\%$ throughput, STDP costs $\sim\!6\%$, and CAEL costs
$<\!0.5\%$. The full pipeline therefore runs at $1.06\times$ the
LIF-only floor while preserving all three formal guarantees, which is
the practical operating point because the system's value comes from the
\emph{joint} cognitive--reproducibility property rather than any single
component in isolation.

% ────────────────────────────────────────────────────────────────────────────
\subsection{Full-Loop Demo}
\label{sec:demo}

We walk through an end-to-end instance of the full pipeline on the embodied
navigation task from Section~\ref{sec:embodied}, demonstrating all three
contributions in sequence through the shipped HoloScript navigation
harness.  The supplementary verifier re-runs the same full loop and
asserts the goal, trace, and hash-chain facts used here.

\begin{enumerate}[leftmargin=*,topsep=2pt]
  \item \textbf{Scene instantiation.}  A 10-object navigation scene is
    defined in \texttt{.holo} composition format and instantiated by the
    \holoscript{} runtime.  The SNN agent brain is loaded as an
    \texttt{@npc} \texttt{@llm\_agent} object with 262{,}144 LIF neurons
    distributed across three cortical layers.
  \item \textbf{LIF forward pass.}  The WGSL compute kernel dispatches
    $\lceil 262{,}144 / 256 \rceil = 1{,}024$ workgroups per timestep.
    On the Intel UHD gen-12lp integrated GPU, peak throughput reaches
    $\mathbf{1{,}481.0}$\,M\,neurons/s (Table~\ref{tab:throughput}),
    comfortably above the 60\,Hz interactive budget.
    Chrome WebGPU is the dispatch target; no plugins, no server required.
  \item \textbf{Tropical APSP for connectivity analysis.}  Between episodes
    the agent queries the SNN topology graph ($N = 2{,}048$ nodes) via
    tropical min-plus GEMM squaring.  On the NVIDIA RTX 3060 Laptop
    (discrete Ampere), the full APSP closure completes in
    $\mathbf{740.9}$\,ms at $255.066$\,GFLOP/s---a
    $\mathbf{27.60\times}$ speedup over the integrated baseline
    (Table~\ref{tab:tropical-crossgpu}).
  \item \textbf{CAEL provenance.}  Every perception--cognition--action
    cycle emits a \cael{} trace entry: the spike-rate vector, the
    tropical activation output, and the action label are FNV-1a
    hash-chained together.  The checked demo covers the complete
    200-tick navigation episode with four CAEL records per tick,
    producing an 800-entry tamper-evident audit log.
  \item \textbf{Reproducibility anchor.}  The supplementary verifier
    runs the HoloScript harness and fails unless the goal is reached at
    tick~38, the CAEL trace has 800 entries, and the hash chain verifies.
    Benchmark snapshots and RTX runs remain separately OTS-anchored in
    the evidence memos cited in \S\ref{sec:evaluation}.
\end{enumerate}

\paragraph{Provenance chain composition.}
The CAEL chain is not hand-waved hashing; it has a fixed inductive
structure. Let $h_{-1} = \text{FNV-1a}(\text{SHA-256}(\text{compiled
scene}))$ be the genesis link. For each timestep $t \in [0, T)$:
\[
  h_t = \text{FNV-1a}\!\left(h_{t-1} \,\Vert\, \mathbf{s}_t \,\Vert\, \mathbf{a}_t \,\Vert\, \ell_t\right)
\]
where $\mathbf{s}_t$ is the spike-rate vector emitted by the LIF
kernel, $\mathbf{a}_t$ is the tropical-activation output passed to
the policy head, $\ell_t$ is the action label, and $\Vert$ is byte
concatenation in canonical little-endian. The terminal head $h_{T}$
at episode end ($T = 200$ ticks in the canonical demo harness) is what
the verifier recomputes. Tampering with any single
$\mathbf{s}_t$, $\mathbf{a}_t$, or $\ell_t$ in the log invalidates
$h_T$ and is detected on replay.

\paragraph{Reproduction protocol.}
An independent auditor verifies the run as follows:
(a)~obtain the supplementary bundle under
\texttt{research/paper-2-demo/};
(b)~run \texttt{node reproduce.mjs -{}-check}, which re-executes
\texttt{packages/engine/src/simulation/\_\_tests\_\_/paper-snn-navigation.test.ts};
(c)~assert that the verifier reports goal reached at tick~38, 800 CAEL
entries, hash-chain validity, and a passing Vitest file.  A divergence
localises either to the HoloScript harness output or to the verifier's
pattern checks, both of which are visible in the command output.

The complete verifier and a local hardware verification memo are included
in \texttt{research/paper-2-demo/} (D.011 full-loop evidence).
\measuredFrom{research/paper-2-demo/reproduce.mjs}%
\measuredFrom{research/paper-2-demo/FULL_LOOP_VERIFICATION_2026-05-10.md}%

\section{Related Work}
\label{sec:related}

\paragraph{SNN Simulation Frameworks.}
NEST~\cite{gewaltig2007nest} is the gold standard for large-scale SNN simulation
($>10^6$ neurons) but requires a Linux cluster.
BrainCog~\cite{tang2024braincog} provides a PyTorch-based SNN framework with
embodied AI integration (iCub, Robosim) but requires CUDA GPUs.
snnTorch~\cite{eshraghian2023training} focuses on training SNNs with surrogate
gradients---a complementary concern to our GPU inference focus.
SpikingJelly~\cite{fang2023spikingjelly} offers CUDA-accelerated SNN simulation
with neuromorphic dataset support.
None of these run in a web browser.

\paragraph{Web-Based Neural Networks.}
TensorFlow.js~\cite{smilkov2019tensorflow} supports DNNs on WebGPU but does
not implement spiking dynamics.
ONNX Runtime Web uses WebGPU for standard neural network inference.
Neither provides spike encoding/decoding, STDP, or LIF neuron models.
To our knowledge, \snnwebgpu{} is the first SNN implementation targeting WebGPU.

\paragraph{Neuromorphic Hardware.}
Intel Loihi 2~\cite{davies2018loihi} and IBM TrueNorth~\cite{merolla2014truenorth}
implement SNNs in custom silicon with $>10^6$ neurons and $< 1$W power.
SpiNNaker~\cite{furber2014spinnaker} uses ARM cores for flexible neuron models.
These provide superior energy efficiency but require specialized hardware
procurement and toolchains---the antithesis of our zero-installation approach.

\paragraph{Spiking World Models.}
Capone and Paolucci~\cite{capone2024dreaming} (Sci.\ Rep.\ 2024) demonstrate
biologically plausible model-based reinforcement learning in recurrent spiking
networks, where ``dreaming''---generating simulated experiences inside a
spike-based world model---accelerates skill acquisition without offline replay
buffers.
Our work complements this by showing that spike-based world-model execution can
run end-to-end in a browser with provenance over every spike, rather than
requiring HPC infrastructure or specialized neuromorphic hardware.

\paragraph{Tropical Algebra in Neural Networks.}
The connection between tropical geometry and neural networks has been explored
theoretically~\cite{zhang2018tropical, alfarra2022decision}, showing that ReLU
networks partition input space into tropical hypersurfaces.
Maragos et al.~\cite{maragos2021tropical} provide a comprehensive survey of
tropical algebra in signal processing and machine learning.
Our contribution is the \emph{implementation}: a GPU-accelerated tropical bridge
that connects SNN rate codes to ReLU activations in a working system with
empirical benchmarks.

\paragraph{Provenance in Simulation.}
Simulation provenance has been explored in scientific workflows~\cite{simmhan2005survey}
and digital twins~\cite{tao2019digital}, but always at the simulation level---not
at the individual neural decision level.
Our \cael{} system provides per-spike-train, per-action hash-chain provenance,
enabling counterfactual analysis (``what if the agent had chosen differently?'')
with cryptographic integrity guarantees.

% ============================================================================
\section{Discussion}
\label{sec:discussion}

\paragraph{Limitations.}
Our LIF model does not capture the full richness of biological neurons
(e.g., dendritic computation, neuromodulation, short-term plasticity).
The STDP rule is a simplified Hebbian variant; biologically accurate STDP
requires pre/post spike timing windows that our current implementation
does not track across timesteps.
WebGPU is not yet universally available---Safari support arrived in 2024,
but older browsers and some mobile devices lack support.
The max-plus tropical bridge (Theorem~\ref{thm:bridge}) holds in the
suprathreshold steady-state regime; transient dynamics and
sub-threshold oscillations break the linear approximation.

\paragraph{Broader Impact.}
Browser-native neuromorphic cognition democratizes access to embodied AI
research.
A student can open a URL and run an SNN-driven agent in a physics simulation
without installing Python, CUDA, or requesting cluster time.
The \cael{} provenance system provides a foundation for auditable AI
decision-making, which is increasingly important as autonomous systems
enter safety-critical domains.
The tropical bridge offers a theoretical unification that could inform
curriculum design for teaching both biological and artificial neural
computation.

\paragraph{Future Work.}
\label{sec:future}
Item~(1) below was originally listed as future work but is now closed:
the STDP-learned navigation result is reported in
Section~\ref{sec:stdp-nav} (Table~\ref{tab:stdp-nav-experiment}).
Remaining next steps:
(1) [\emph{closed; see Section~\ref{sec:stdp-nav}}] enabling STDP learning
in the navigation task to optimize path efficiency beyond the baseline
20\%; the soft-bounded Hebbian rule with a supervised teacher trains a
random network to a near-optimal $39$-tick / $29.97$-unit path in 8
episodes (CPU reference, deterministic, $60$\,ms wall);
(2) implementing biologically accurate STDP with exponential timing windows;
(3) multi-agent \cael{} traces with causal ordering across agents;
(4) integration with neuromorphic hardware (Loihi 2) for power efficiency
comparisons;
(5) scaling to $10^6$ neurons via hierarchical workgroup decomposition
and shared memory tiling;
(6) connecting the tropical graph algorithms to network architecture
search---finding optimal SNN topologies via shortest-path analysis
of information flow.

\subsection{Remaining Work for Camera-Ready}
\label{sec:remaining-work}
\begin{itemize}
  \item In-repo SHA-256 vs FNV-1a microbenchmark on the canonical host:
    \emph{shipped} as
    \texttt{packages/engine/src/simulation/\_\_tests\_\_/fnv1a-vs-sha256.bench.test.ts}
    (HoloScript). Pure-JS SHA-256 measured at $\sim$9--24$\times$
    per-hash cost vs FNV-1a across paper-3 §7.5 scale ladder
    (60/300/1{,}200/6{,}000 floats plus 500\,B--500\,KB payloads);
    validated against RFC~6234 §B.1--2 test vectors and cross-checked
    against Node native \texttt{crypto.createHash('sha256')} on 9
    random-input sizes spanning single-block / block-boundary /
    multi-block cases.
\end{itemize}

% ============================================================================
\section{Conclusion}
\label{sec:conclusion}

We have presented \snnwebgpu{}, the first browser-native spiking neural network
system running on WebGPU compute shaders, integrated with physics simulation
via hash-chain provenance.
The system implements LIF neurons, STDP learning, 7 spike encoding/decoding
schemes, and tropical algebra operations entirely in WGSL shaders dispatched
from a standard web browser.
On the integrated baseline (Intel UHD Graphics gen-12lp), the LIF kernel
sustains $1.48 \times 10^9$ neuron-timesteps/s at $N = 262{,}144$ with
the browser-native WGSL path and no installation. The 2026-05-06 strict
adapter RTX~3060 rerun reaches $5.89 \times 10^9$ neuron-timesteps/s at
$N = 1{,}048{,}576$.
We established a formal connection between SNN rate coding and ReLU activations
through the tropical max-plus semiring, implemented as a GPU-accelerated bridge.
Our tropical min-plus GEMM achieves $20.64\times$ speedup over CPU at $N = 1024$
on integrated graphics (peak $9.88$\,GFLOP/s) and scales to $\mathbf{255.066}$
\textbf{GFLOP/s} at $N = 2048$ on the discrete GPU---a $27.60\times$ cross-GPU
speedup at the largest size---enabling online APSP analysis of SNN connectivity
graphs during simulation.
Every neural decision in the embodied agent loop is committed to a tamper-evident
hash-chain trace at a cost of $3.40\;\mu$s per entry median ($3.74\;\mu$s p99;
$\sim\!0.082\%$ / $0.090\%$ of a 60\,Hz frame budget at median/p99 when charging four trace entries per tick), producing the first compute substrate to emit auditable, replayable traces of neuromorphic
cognition at hardware throughput in physics simulation.

Browser-native SNN is deployable on integrated consumer graphics today
($1.48 \times 10^9$ neuron-timesteps/s baseline), with the same codebase,
shader, and WebGPU dispatch path scaling to a verified RTX~3060 laptop run
at $5.89 \times 10^9$ neuron-timesteps/s.
In an embodied navigation task, a 128-neuron SNN agent reaches its goal in
38 ticks (1.9\,s simulation time) at a mean tick time of 0.44\,ms---a
$\sim$2{,}270\,Hz agent loop rate with median/mean ticks $\approx 2.1\% / 2.7\%$ of a 60\,Hz frame
budget---demonstrating that goal-directed behavior emerges from asymmetric
input encoding (obstacle inhibition, goal excitation) without explicit path
planning.

The complete system---7 WGSL shaders, 16 compute entry points, 205 Vitest tests
in \texttt{@holoscript/snn-webgpu} (2026-04-17), and a multi-layer SNN orchestrator with STDP
learning---is open-source and runs in any WebGPU-capable browser with zero
installation.

More fundamentally, this paper establishes the \emph{determinism receipt} that
makes NN-primary runtime stacks architecturally honest: a spiking neural network
engine whose every decision is cryptographically traceable, reproducibly
auditable, and formally bounded by a SimulationContract verifier.
Paper~2 is therefore the credential paper for an inverted runtime in which
neuromorphic computation is the primary executor and CPU-side verification is the
backup referee---generalizing speculative decoding from token-stream to
operation-stream and opening the path to always-on embodied agents at
single-digit milliwatt power.

% ============================================================================
\section*{Reproducibility Statement}

All source code is available as the \texttt{@holoscript/snn-webgpu} package
within the HoloScript repository.
The benchmark harness (\texttt{tropical-shortest-paths.benchmark.test.ts})
is self-contained and runnable via \texttt{pnpm test}.
WGSL shader sources are provided in the supplementary materials.
The \cael{} trace format is documented in the
\texttt{CAELTrace.ts} module with full serialization/deserialization
and hash-chain verification functions.

% ============================================================================
\section*{Acknowledgments}

This work was developed within the HoloScript ecosystem by Infinitus Systems.
We thank the W3C WebGPU Community Group for the compute shader specification
that made this work possible.

% ============================================================================
\iffalse
% Legacy inline bibliography preserved for audit trail.
% Wrapped 2026-05-04 per bib-migration ruling.
% Active bibliography: \bibliography{holoscript} at end of document.
\bibliographystyle{plain}

\begin{thebibliography}{20}

\bibitem{maass1997networks}
W.~Maass.
\newblock Networks of spiking neurons: The third generation of neural network
  models.
\newblock {\em Neural Networks}, 10(9):1659--1671, 1997.

\bibitem{gerstner2002spiking}
W.~Gerstner and W.~M. Kistler.
\newblock {\em Spiking Neuron Models: Single Neurons, Populations, Plasticity}.
\newblock Cambridge University Press, 2002.

\bibitem{tang2024braincog}
H.~Tang, Y.~Zeng, et~al.
\newblock BrainCog: A spiking neural network based brain-inspired cognitive
  intelligence engine for brain-inspired AI and brain simulation.
\newblock {\em Patterns}, 4(8):100789, 2023.

\bibitem{gewaltig2007nest}
M.-O. Gewaltig and M.~Diesmann.
\newblock NEST (NEural Simulation Tool).
\newblock {\em Scholarpedia}, 2(4):1430, 2007.

\bibitem{davies2018loihi}
M.~Davies, N.~Srinivasa, et~al.
\newblock Loihi: A neuromorphic manycore processor with on-chip learning.
\newblock {\em IEEE Micro}, 38(1):82--99, 2018.

\bibitem{merolla2014truenorth}
P.~A. Merolla, J.~V. Arthur, et~al.
\newblock A million spiking-neuron integrated circuit with a scalable
  communication network and interface.
\newblock {\em Science}, 345(6197):668--673, 2014.

\bibitem{webgpu2024spec}
W3C.
\newblock {WebGPU} specification.
\newblock W3C Working Draft, 2024.
\newblock \url{https://www.w3.org/TR/webgpu/}.

\bibitem{smilkov2019tensorflow}
D.~Smilkov, N.~Thorat, et~al.
\newblock TensorFlow.js: Machine learning for the web and beyond.
\newblock In {\em Proceedings of Machine Learning and Systems (MLSys)}, 2019.

\bibitem{maclagan2015introduction}
D.~Maclagan and B.~Sturmfels.
\newblock {\em Introduction to Tropical Geometry}.
\newblock American Mathematical Society, 2015.

\bibitem{holoscript2026cael}
J.~Krzywoszyja.
\newblock {CAEL}: Causal Agent-Environment Loop with hash-chain provenance.
\newblock Technical report, Infinitus Systems, 2026.

\bibitem{glorot2011deep}
X.~Glorot, A.~Bordes, and Y.~Bengio.
\newblock Deep sparse rectifier neural networks.
\newblock In {\em Proceedings of AISTATS}, pages 315--323, 2011.

\bibitem{blelloch1990prefix}
G.~E. Blelloch.
\newblock Prefix sums and their applications.
\newblock Technical Report CMU-CS-90-190, Carnegie Mellon University, 1990.

\bibitem{capone2024dreaming}
C.~Capone and P.~S. Paolucci.
\newblock Towards biologically plausible model-based reinforcement learning in
  recurrent spiking networks by dreaming new experiences.
\newblock {\em Scientific Reports}, 14(1):14656, 2024.
\newblock DOI: 10.1038/s41598-024-65631-y. Replaces hallucinated
  \texttt{zhang2025spiking} stub (2026-05-11).

\bibitem{zhang2018tropical}
L.~Zhang, G.~Naitzat, and L.~H. Lim.
\newblock Tropical geometry of deep neural networks.
\newblock In {\em Proceedings of ICML}, pages 5824--5832, 2018.

\bibitem{alfarra2022decision}
M.~Alfarra, A.~Bibi, et~al.
\newblock On the decision boundaries of neural networks: A tropical geometry
  perspective.
\newblock {\em IEEE Transactions on Pattern Analysis and Machine Intelligence},
  45(10):12312--12324, 2022.

\bibitem{maragos2021tropical}
P.~Maragos, V.~Charisopoulos, and E.~Theodosis.
\newblock Tropical geometry and machine learning.
\newblock {\em Proceedings of the IEEE}, 109(5):728--755, 2021.

\bibitem{simmhan2005survey}
Y.~L. Simmhan, B.~Plale, and D.~Gannon.
\newblock A survey of data provenance in e-science.
\newblock {\em ACM SIGMOD Record}, 34(3):31--36, 2005.

\bibitem{tao2019digital}
F.~Tao, H.~Zhang, et~al.
\newblock Digital twin in industry: State-of-the-art.
\newblock {\em IEEE Transactions on Industrial Informatics}, 15(4):2405--2415,
  2019.

\bibitem{eshraghian2023training}
J.~K. Eshraghian, M.~Ward, et~al.
\newblock Training spiking neural networks using lessons from deep learning.
\newblock {\em Proceedings of the IEEE}, 111(9):1016--1054, 2023.

\bibitem{fang2023spikingjelly}
W.~Fang, Y.~Chen, et~al.
\newblock SpikingJelly: An open-source machine learning infrastructure platform
  for spike-based intelligence.
\newblock {\em Proceedings of the IEEE}, 111(9):1055--1078, 2023.

\bibitem{furber2014spinnaker}
S.~B. Furber, F.~Galluppi, S.~Temple, and L.~A. Plana.
\newblock The {SpiNNaker} project.
\newblock {\em Proceedings of the IEEE}, 102(5):652--665, 2014.

\end{thebibliography}
\fi

% ============================================================================
\appendix

\section{Default LIF Parameters}
\label{app:lif-params}

\begin{table}[h]
\centering
\caption{Default LIF parameters used throughout experiments, based on standard
  cortical neuron models~\cite{gerstner2002spiking}.}
\label{tab:lif-params}
\small
\begin{tabular}{@{}llrl@{}}
\toprule
\textbf{Parameter} & \textbf{Symbol} & \textbf{Value} & \textbf{Unit} \\
\midrule
Membrane time constant   & $\tau$           & 20.0  & ms \\
Spike threshold          & $V_\theta$       & $-55.0$ & mV \\
Reset voltage            & $V_\mathrm{reset}$  & $-75.0$ & mV \\
Resting potential        & $V_\mathrm{rest}$   & $-65.0$ & mV \\
Timestep                 & $\Delta t$       & 1.0   & ms \\
Refractory period        & $t_\mathrm{ref}$ & 2.0   & ms \\
\bottomrule
\end{tabular}
\end{table}

\section{Full WGSL Shader Listing}
\label{app:shaders}

The complete WGSL sources for all 7 shader modules (comprising 16 compute entry points)
are provided in the supplementary materials.
Key characteristics:
\begin{itemize}
\item All kernels use workgroup size 256 (1D) except \texttt{tropical\_min\_plus\_gemm}
  which uses $16 \times 16$ (2D).
\item Infinity is encoded as \texttt{1e30} (the \texttt{TROPICAL\_INF} sentinel),
  safely below the IEEE 754 float32 infinity to avoid NaN propagation in
  \texttt{min}/\texttt{max} operations.
\item The PCG hash function in \texttt{spike-encode.wgsl} (period $2^{32}$) provides
  uniform random numbers for rate coding without requiring a global RNG state buffer.
\item The Blelloch scan (\texttt{blelloch-scan-sort.wgsl}) uses only
  \texttt{workgroupBarrier()} for synchronization---no atomics on global memory---ensuring
  compatibility with Apple M-series and Intel Arc GPUs where global atomics can deadlock.
\end{itemize}

\section{Tropical Semiring Properties}
\label{app:tropical}

For completeness, we state the semiring axioms satisfied by our implementation:

\begin{itemize}
\item \textbf{Max-plus} $(\mathbb{R} \cup \{-\infty\}, \tropplus = \max, \tropmult = +)$:
  \begin{itemize}
    \item Identity for $\tropplus$: $-\infty$ (tropical zero)
    \item Identity for $\tropmult$: $0$ (tropical one)
    \item $\tropplus$ is commutative, associative, idempotent
    \item $\tropmult$ distributes over $\tropplus$: $a + \max(b,c) = \max(a+b, a+c)$
  \end{itemize}

\item \textbf{Min-plus} $(\mathbb{R} \cup \{+\infty\}, \tropplus = \min, \tropmult = +)$:
  \begin{itemize}
    \item Identity for $\tropplus$: $+\infty$ (tropical zero, encoded as \texttt{1e30})
    \item Identity for $\tropmult$: $0$ (tropical one)
    \item Tropical matrix product: $(A \tropmult B)_{ij} = \min_k (A_{ik} + B_{kj})$
    \item Repeated squaring $A^{\tropmult \lceil \log_2 N \rceil}$ yields APSP distances
  \end{itemize}
\end{itemize}

The \texttt{TropicalActivationTrait} bridges the two semirings:
max-plus variant computes $\alpha \cdot \max(0, x - \theta)$ (ReLU);
min-plus variant computes $\alpha \cdot \min(0, x - \theta)$ (cost dual).
Both are dispatched as a single GPU kernel with a variant flag.

\section{Broader Impact}
\label{app:broader-impact}

This work accelerates embodied spiking neural networks (SNNs) in consumer browsers via WebGPU, lowering the barrier to entry for neuromorphic research and edge-based agent cognition. While democratizing access to neuro-symbolic simulation is broadly positive (enabling decentralized robotics research and privacy-preserving local computation), there are potential risks. The deterministic provenance chains (\cael{}) provide auditability, which mitigates adversarial manipulation, but open-source GPU-accelerated neuromorphic agents could be deployed in autonomous systems lacking adequate physical safeguards. We encourage the continued pairing of contracted physics (which enforce safety bounds at inference) with any downstream usage of this acceleration framework.

\bibliographystyle{plain}
\bibliography{holoscript}

\end{document}
